---
title: "Basic Questions For Curated and Collated SCFA Data"
author: "Alex Yang"
date: "9/21/2020"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
#Loading Appropriate Packages 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('dplyr')
library('tidyverse')
library('ggplot2')
library('scales')
library('ggpubr')
```

```{r}
#Always Account for Working Directory
getwd()
```

#-------------------------------------------------------------------------------

#SCFA vs. PH
##SCFA vs. PH for Week 1 and Week 3 (Participants Averaged)
```{r}
#Reading in Fecal Sample Data
fullfecalsample <- read.csv("./fecalsample_a.csv")

fecal_clean <- fullfecalsample %>%
  na.omit() %>%
  filter(Week == "week1" | Week == "week3") %>%
  mutate(scfa_sum = rowSums(.[6:8])) 
#Preparing Data for Scatterplot Creation 
fecal_clean$pH = as.numeric(fecal_clean$pH)
fecal_clean$scfa_sum = as.numeric(fecal_clean$scfa_sum)

#Creating Scatterplot for SCFA vs. PH
ggplot(fecal_clean, aes(x=pH, y=scfa_sum)) +
  geom_point(alpha=0.3) +
  ggtitle("SCFA vs. PH for All Weeks") +
  labs(y= "Total SCFA", x = "Average PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH without Points
ggplot(fecal_clean, aes(x=pH, y=scfa_sum)) +
  geom_smooth() +
  ggtitle("SCFA vs. PH for All Weeks") +
  labs(y= "Total SCFA", x = "Average PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH with Points
ggplot(fecal_clean, aes(x=pH, y=scfa_sum)) +
  geom_point() +
  geom_smooth() +
  ggtitle("SCFA vs. PH for All Weeks") +
  labs(y= "Total SCFA", x = "Average PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Correlation Test
cor(as.vector(fecal_clean$pH), as.vector(fecal_clean$scfa_sum),
    method = "pearson") #Returns -0.1326739. Low negative correlation 

#Significance Statistics
qqplot(as.vector(fecal_clean$pH), as.vector(fecal_clean$scfa_sum))
bartlett.test(scfa_sum ~ Week, data = fecal_clean)
cor.test(as.vector(fecal_clean$pH), as.vector(fecal_clean$scfa_sum), 
         alternative = "less", method = 'pearson')
```

##SCFA vs. PH for Week 1 and Week 3 (Participants Unaveraged)
```{r}
#Reading in Fecal Sample Data
fullfecalsample_unaveraged <- read.csv("./fecalsample_ua.csv")
fecal_clean_ua <- fullfecalsample_unaveraged %>%
  na.omit() %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH"),
         starts_with("Acetate_mmol_kg"),
         starts_with("Butyrate_mmol_kg"),
         starts_with("Propionate_mmol_kg")) %>%
  filter(Week == "week1" | Week == "week3") %>%
  mutate(scfa_sum = rowSums(.[4:6])) 

#Preparing Data for Scatterplot Creation 
fecal_clean_ua$pH = as.numeric(fecal_clean_ua$pH)
fecal_clean_ua$scfa_sum = as.numeric(fecal_clean_ua$scfa_sum)

#Creating Scatterplot for SCFA vs. PH
ggplot(fecal_clean_ua, aes(x=pH, y=scfa_sum)) +
  geom_point(alpha=0.3) +
  ggtitle("SCFA vs. PH for All Weeks") +
  labs(y= "Total SCFA", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH without Points
ggplot(fecal_clean_ua, aes(x=pH, y=scfa_sum)) +
  geom_smooth() +
  ggtitle("SCFA vs. PH for All Weeks") +
  labs(y= "Total SCFA", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH with Points
ggplot(fecal_clean_ua, aes(x=pH, y=scfa_sum)) +
  geom_point() +
  geom_smooth() +
  ggtitle("SCFA vs. PH for All Weeks") +
  labs(y= "Total SCFA", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Correlation Test
cor(as.vector(fecal_clean_ua$pH), as.vector(fecal_clean_ua$scfa_sum),
    method = "pearson") #Returns -0.1765633. Low negative correlation

#Significance Statistics
qqplot(as.vector(fecal_clean_ua$pH), as.vector(fecal_clean_ua$scfa_sum))
bartlett.test(scfa_sum ~ Week, data = fecal_clean_ua)
cor.test(as.vector(fecal_clean_ua$pH), as.vector(fecal_clean_ua$scfa_sum), 
         alternative = "less", method = 'pearson')
```

##SCFA vs. PH for Week 1 Only (Participants Averaged)
```{r}
#Reading in Fecal Sample Data
fullfecalsample <- read.csv("./fecalsample_a.csv")
fecal_clean1 <- fullfecalsample %>%
  na.omit() %>%
  filter(Week == "week1") %>%
  mutate(scfa_sum = rowSums(.[6:8])) 

#Preparing Data for Scatterplot Creation 
fecal_clean1$pH = as.numeric(fecal_clean1$pH)
fecal_clean1$scfa_sum = as.numeric(fecal_clean1$scfa_sum)

#Creating Scatterplot for SCFA vs. PH
ggplot(fecal_clean1, aes(x=pH, y=scfa_sum)) +
  geom_point(alpha=0.3) +
  ggtitle("SCFA vs. PH for Week 1") +
  labs(y= "Total SCFA", x = "Average PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH without Points
ggplot(fecal_clean1, aes(x=pH, y=scfa_sum)) +
  geom_smooth() +
  ggtitle("SCFA vs. PH for Week 1") +
  labs(y= "Total SCFA", x = "Average PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH with Points
ggplot(fecal_clean1, aes(x=pH, y=scfa_sum)) +
  geom_point() +
  geom_smooth() +
  ggtitle("SCFA vs. PH for Week 1") +
  labs(y= "Total SCFA", x = "Average PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Correlation Test
cor(as.vector(fecal_clean1$pH), as.vector(fecal_clean1$scfa_sum),
    method = "pearson") #Returns -0.1444287. Negative correlation

#Significance Statistics
cor.test(as.vector(fecal_clean1$pH), as.vector(fecal_clean1$scfa_sum), 
         alternative = "less", method = 'pearson')
```

##SCFA vs. PH for Week 1 Only (Participants Unaveraged)
```{r}
#Reading in Fecal Sample Data
fullfecalsample_unaveraged <- read.csv("./fecalsample_ua.csv")
fecal_clean_ua <- fullfecalsample_unaveraged %>%
  na.omit() %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH"),
         starts_with("Acetate_mmol_kg"),
         starts_with("Butyrate_mmol_kg"),
         starts_with("Propionate_mmol_kg")) %>%
  filter(Week == "week1") %>%
  mutate(scfa_sum = rowSums(.[4:6]))

#Preparing Data for Scatterplot Creation 
fecal_clean_ua$pH = as.numeric(fecal_clean_ua$pH)
fecal_clean_ua$scfa_sum = as.numeric(fecal_clean_ua$scfa_sum)

#Creating Scatterplot for SCFA vs. PH
ggplot(fecal_clean_ua, aes(x=pH, y=scfa_sum)) +
  geom_point(alpha=0.3) +
  ggtitle("SCFA vs. PH for All Weeks") +
  labs(y= "Total SCFA", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH without Points
ggplot(fecal_clean_ua, aes(x=pH, y=scfa_sum)) +
  geom_smooth() +
  ggtitle("SCFA vs. PH for All Weeks") +
  labs(y= "Total SCFA", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH with Points
ggplot(fecal_clean_ua, aes(x=pH, y=scfa_sum)) +
  geom_point() +
  geom_smooth() +
  ggtitle("SCFA vs. PH for All Weeks") +
  labs(y= "Total SCFA", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Correlation Test
cor(as.vector(fecal_clean_ua$pH), as.vector(fecal_clean_ua$scfa_sum),
    method = "pearson") #Returns -0.1765633. Low negative correlation

#Significance Statistics
qqplot(as.vector(fecal_clean_ua$pH), as.vector(fecal_clean_ua$scfa_sum))
cor.test(as.vector(fecal_clean_ua$pH), as.vector(fecal_clean_ua$scfa_sum), 
         alternative = "less", method = 'pearson')
```


##SCFA vs. PH for Week 3 Only (Participants Averaged)
```{r}
#Reading in Fecal Sample Data
fullfecalsample <- read.csv("./fecalsample_a.csv")
fecal_clean3 <- fullfecalsample %>%
  na.omit() %>%
  filter(Week == "week3") %>%
  mutate(scfa_sum = rowSums(.[6:8]))

#Preparing Data for Scatterplot Creation 
fecal_clean3$pH = as.numeric(fecal_clean3$pH)
fecal_clean3$scfa_sum = as.numeric(fecal_clean3$scfa_sum)

#Creating Scatterplot for SCFA vs. PH
ggplot(fecal_clean3, aes(x=pH, y=scfa_sum)) +
  geom_point(alpha=0.3) +
  ggtitle("SCFA vs. PH for Week 3") +
  labs(y= "Total SCFA", x = "Average PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH without Points
ggplot(fecal_clean3, aes(x=pH, y=scfa_sum)) +
  geom_smooth() +
  ggtitle("SCFA vs. PH for Week 3") +
  labs(y= "Total SCFA", x = "Average PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH with Points
ggplot(fecal_clean3, aes(x=pH, y=scfa_sum)) +
  geom_point() +
  geom_smooth() +
  ggtitle("SCFA vs. PH for Week 3") +
  labs(y= "Total SCFA", x = "Average PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Correlation Test
cor(as.vector(fecal_clean3$pH), as.vector(fecal_clean3$scfa_sum),
    method = "pearson") #Returns -0.1104008. Very low negative correlation.

#Significance Statistics
cor.test(as.vector(fecal_clean3$pH), as.vector(fecal_clean3$scfa_sum), 
         alternative = "less", method = 'pearson')
```

##SCFA vs. PH for Week 3 Only (Participants Unaveraged)
```{r}
#Reading in Fecal Sample Data
fullfecalsample_unaveraged <- read.csv("./fecalsample_ua.csv")
fecal_clean_ua <- fullfecalsample_unaveraged %>%
  na.omit() %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH"),
         starts_with("Acetate_mmol_kg"),
         starts_with("Butyrate_mmol_kg"),
         starts_with("Propionate_mmol_kg")) %>%
  filter(Week == "week3") %>%
  mutate(scfa_sum = rowSums(.[4:6]))

#Preparing Data for Scatterplot Creation 
fecal_clean_ua$pH = as.numeric(fecal_clean_ua$pH)
fecal_clean_ua$scfa_sum = as.numeric(fecal_clean_ua$scfa_sum)

#Creating Scatterplot for SCFA vs. PH
ggplot(fecal_clean_ua, aes(x=pH, y=scfa_sum)) +
  geom_point(alpha=0.3) +
  ggtitle("SCFA vs. PH for All Weeks") +
  labs(y= "Total SCFA", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH without Points
ggplot(fecal_clean_ua, aes(x=pH, y=scfa_sum)) +
  geom_smooth() +
  ggtitle("SCFA vs. PH for All Weeks") +
  labs(y= "Total SCFA", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH with Points
ggplot(fecal_clean_ua, aes(x=pH, y=scfa_sum)) +
  geom_point() +
  geom_smooth() +
  ggtitle("SCFA vs. PH for All Weeks") +
  labs(y= "Total SCFA", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Correlation Test
cor(as.vector(fecal_clean_ua$pH), as.vector(fecal_clean_ua$scfa_sum),
    method = "pearson") #Returns -0.1765633. Low negative correlation

#Significance Statistics
qqplot(as.vector(fecal_clean_ua$pH), as.vector(fecal_clean_ua$scfa_sum))
cor.test(as.vector(fecal_clean_ua$pH), as.vector(fecal_clean_ua$scfa_sum), 
         alternative = "less", method = 'pearson')
```

#-------------------------------------------------------------------------------

#BRMPS 2x Daily Cohort 
##Acetate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
fullfecalsample <- read.csv("./fecalsample_a.csv")
View(fullfecalsample)
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")
View(participantsupps1)

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "BRMPS 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_acetate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_acetate,participantsupps1,by="Participant") %>%
  rename(acetate.wk1 = Acetate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("acetate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_acetate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate")) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_acetate,participantsupps1,by="Participant") %>%
  rename(acetate.wk3 = Acetate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("acetate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
acetate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
acetate_brmps_diff <- acetate_brmps %>%
  mutate(acetate_diff = acetate.wk3 - acetate.wk1) %>%
  arrange(by = acetate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
acetate_wk1_vec <- as.vector(acetate_brmps$acetate.wk1)
acetate_wk3_vec <- as.vector(acetate_brmps$acetate.wk3)
paired_acetate_df <- data.frame(Week1 = acetate_wk1_vec, Week3 = acetate_wk3_vec)
acetatepairplot <- ggpaired(paired_acetate_df, cond1 = "Week1", cond2 = "Week3")
acetatepairplot$layers <- acetatepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(acetate_brmps$acetate.wk1), 
       as.vector(acetate_brmps$acetate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(acetate_brmps_diff)
acetatepairplot
acebrmps1x_barplot <- barplot(acetate_brmps_diff$acetate_diff, ylab = "Participants", xlab = "Acet. Conc. Change (Week 3 - Week 1)", 
        main = "BRMPS 2x Daily Acetate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(acetate_brmps_diff)
```

##Acetate Proportion in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "BRMPS 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_acetateprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week1") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(acetate_prop1 = Acetate_mmol_kg / scfa_sum) 
View(fullfecalsample)

fecal1_brmpsprop <- merge(fecal1_acetateprop,participantsupps1,by="Participant") %>%
  rename(acetate_prop_wk1 = acetate_prop1) %>%
  select(starts_with("Participant"),
         starts_with("acetate_prop_wk1"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_acetateprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week3") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(acetate_prop3 = Acetate_mmol_kg / scfa_sum) 

fecal3_brmpsprop <- merge(fecal3_acetateprop,participantsupps1,by="Participant") %>%
  rename(acetate_prop_wk3 = acetate_prop3) %>%
  select(starts_with("Participant"),
         starts_with("acetate_prop_wk3"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
acetate_brmpsprop <- merge(fecal1_brmpsprop,fecal3_brmpsprop,
                           by="Participant") %>%
                     select(-ends_with("x"), -ends_with("y"))

#Appending Differences column to acetate-BRMPS df in form: week3-week1
acetate_brmps_diffprop <- acetate_brmpsprop %>%
  mutate(acetate_prop_diff = acetate_prop_wk3 - acetate_prop_wk1) %>%
  arrange(by = acetate_prop_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
acetate_wk1_vecprop <- as.vector(acetate_brmps_diffprop$acetate_prop_wk1)
acetate_wk3_vecprop <- as.vector(acetate_brmps_diffprop$acetate_prop_wk3)
paired_acetate_dfprop <- data.frame(Week1 = acetate_wk1_vecprop, 
                                    Week3 = acetate_wk3_vecprop)
acetatepairplotprop <- ggpaired(paired_acetate_dfprop, cond1 = "Week1", 
                                cond2 = "Week3")
acetatepairplotprop$layers <- acetatepairplotprop$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(acetate_brmpsprop$acetate_prop_wk1), 
       as.vector(acetate_brmpsprop$acetate_prop_wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(acetate_brmps_diffprop) # p-value = 0.0971. 
acetatepairplotprop
aceprop1x_barplot <- barplot(acetate_brmps_diffprop$acetate_prop_diff, ylab = "Participants", xlab = "Acetate Prop. Change (Week 3 - Week 1)", 
        main = "BRMPS 2x Daily Ace. Prop. Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(acetate_brmps_diffprop)
```

##Butyrate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
fullfecalsample <- read.csv("./fecalsample_a.csv")

#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015") 

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "BRMPS 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_butyrate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Butyrate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_butyrate,participantsupps1,by="Participant") %>%
  rename(butyrate.wk1 = Butyrate_mmol_kg) %>%
  select(starts_with("Participant"),
        starts_with("butyrate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_butyrate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Butyrate")) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_butyrate,participantsupps1,by="Participant") %>%
  rename(butyrate.wk3 = Butyrate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("butyrate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
butyrate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round) 

#Appending Differences column to acetate-BRMPS df in form: week3-week1
butyrate_brmps_diff <- butyrate_brmps %>%
  mutate(butyrate_diff = butyrate.wk3 - butyrate.wk1) %>%
  arrange(by = butyrate_diff)

butyrate_brmps_diff <- butyrate_brmps_diff[-c(86, 87, 88), ]

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
butyrate_wk1_vec <- as.vector(butyrate_brmps$butyrate.wk1)
butyrate_wk3_vec <- as.vector(butyrate_brmps$butyrate.wk3)
paired_butyrate_df <- data.frame(Week1 = butyrate_wk1_vec, Week3 = butyrate_wk3_vec)
butyratepairplot <- ggpaired(paired_butyrate_df, cond1 = "Week1", cond2 = "Week3")
butyratepairplot$layers <- butyratepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(butyrate_brmps_diff$butyrate.wk1), 
       as.vector(butyrate_brmps_diff$butyrate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(butyrate_brmps_diff)
butyratepairplot
butbrmps1x_barplot <- barplot(butyrate_brmps_diff$butyrate_diff, ylab = "Participants", xlab = "Butyr. Conc. Change (Week 3 - Week 1)", 
        main = "BRMPS 2x Daily Butyrate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(butyrate_brmps_diff)
```

##Propionate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "BRMPS 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_propionate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Propionate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_propionate,participantsupps1,by="Participant") %>%
  rename(propionate.wk1 = Propionate_mmol_kg) %>%
  select(starts_with("Participant"),
        starts_with("propionate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_propionate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Propionate")) %>%
  filter(Week == "week3") %>%
  na.omit()

fecal3_brmps <- merge(fecal3_propionate,participantsupps1,by="Participant") %>%
  rename(propionate.wk3 = Propionate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("propionate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
propionate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
propionate_brmps_diff <- propionate_brmps %>%
  mutate(propionate_diff = propionate.wk3 - propionate.wk1) %>%
  arrange(by = propionate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
propionate_wk1_vec <- as.vector(propionate_brmps$propionate.wk1)
propionate_wk3_vec <- as.vector(propionate_brmps$propionate.wk3)
paired_propionate_df <- data.frame(Week1 = propionate_wk1_vec, Week3 = propionate_wk3_vec)
propionatepairplot <- ggpaired(paired_propionate_df, cond1 = "Week1", cond2 = "Week3")
propionatepairplot$layers <- propionatepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(propionate_brmps$propionate.wk1), 
       as.vector(propionate_brmps$propionate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(propionate_brmps_diff)
propionatepairplot
probrmps1x_barplot <- barplot(propionate_brmps_diff$propionate_diff, ylab = "Participants", xlab = "Propio. Conc. Change (Week 3 - Week 1)", 
        main = "BRMPS 1x Daily Propionate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(propionate_brmps_diff)
```
##Prop+But Proportion (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "BRMPS 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_bpprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week1") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(bp_prop1 = (Butyrate_mmol_kg+Propionate_mmol_kg) / scfa_sum) 
#View(fullfecalsample)

fecal1_brmpsprop <- merge(fecal1_bpprop,participantsupps1,by="Participant") %>%
  rename(bp_prop_wk1 = bp_prop1) %>%
  select(starts_with("Participant"),
         starts_with("bp_prop_wk1"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_bpprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week3") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(bp_prop3 = (Butyrate_mmol_kg+Propionate_mmol_kg) / scfa_sum) 

fecal3_brmpsprop <- merge(fecal3_bpprop,participantsupps1,by="Participant") %>%
  rename(bp_prop_wk3 = bp_prop3) %>%
  select(starts_with("Participant"),
         starts_with("bp_prop_wk3"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
bp_brmpsprop <- merge(fecal1_brmpsprop,fecal3_brmpsprop,
                           by="Participant") %>%
                     select(-ends_with("x"), -ends_with("y"))

#Appending Differences column to acetate-BRMPS df in form: week3-week1
bp_brmps_diffprop <- bp_brmpsprop %>%
  mutate(bp_prop_diff = bp_prop_wk3 - bp_prop_wk1) %>%
  arrange(by = bp_prop_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
bp_wk1_vecprop <- as.vector(bp_brmps_diffprop$bp_prop_wk1)
bp_wk3_vecprop <- as.vector(bp_brmps_diffprop$bp_prop_wk3)
paired_bp_dfprop <- data.frame(Week1 = bp_wk1_vecprop, 
                                    Week3 = bp_wk3_vecprop)
bppairplotprop <- ggpaired(paired_bp_dfprop, cond1 = "Week1", 
                                cond2 = "Week3")
bppairplotprop$layers <- bppairplotprop$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(bp_brmpsprop$bp_prop_wk1), 
       as.vector(bp_brmpsprop$bp_prop_wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(bp_brmps_diffprop) # p-value = 0.0971. 
bppairplotprop
bpprop1x_barplot <- barplot(bp_brmps_diffprop$bp_prop_diff, ylab = "Participants", xlab = "BP Prop. Change (Week 3 - Week 1)", 
        main = "BRMPS 2x Daily BP. Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(bp_brmps_diffprop)
```

##Total SCFA Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "BRMPS 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_total <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate")) %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_total,participantsupps1,by="Participant") %>%
  rename(total.wk1 = scfa_sum) %>%
  select(starts_with("Participant"),
        starts_with("total"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_total <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate")) %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_total,participantsupps1,by="Participant") %>%
  rename(total.wk3 = scfa_sum) %>%
  select(starts_with("Participant"),
         starts_with("total"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
total_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
total_brmps_diff <- total_brmps %>%
  mutate(total_diff = total.wk3 - total.wk1) %>%
  arrange(by = total_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
total_wk1_vec <- as.vector(total_brmps$total.wk1)
total_wk3_vec <- as.vector(total_brmps$total.wk3)
paired_total_df <- data.frame(Week1 = total_wk1_vec, Week3 = total_wk3_vec)
totalpairplot <- ggpaired(paired_total_df, cond1 = "Week1", cond2 = "Week3")
totalpairplot$layers <- totalpairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(total_brmps$total.wk1), 
       as.vector(total_brmps$total.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(total_brmps_diff)
totalpairplot
totbrmps1x_barplot <- barplot(total_brmps_diff$total_diff, ylab = "Participants", xlab = "Total Conc. Change (Week 3 - Week 1)", 
        main = "BRMPS 2x Daily Total Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(total_brmps_diff)
```

##pH Change in BRMPS 2x Cohort from Week 1 to Week 3
```{r}
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015") %>%
  rename(Participant = Name) %>%
  filter(Supplement.consumed == "BRMPS 2xdaily 20")

#Making Week 1 BRMPS 2x-pH Data Frame 
ph_brmps_wk1 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week1") %>%
  na.omit() %>%
  rename(ph.wk1 = pH)

#Making Week 3 BRMPS 2x-pH Data Frame 
ph_brmps_wk3 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week3") %>%
  na.omit() %>%
  rename(ph.wk3 = pH)

#Combining Week 1 and Week 3 BRMPS 2x Daily Cohort 
merged_ph_wks <- merge(ph_brmps_wk1, ph_brmps_wk3, by = "Participant") %>%
  select(-starts_with("Week"))
merged_ph_brmps<- merge(merged_ph_wks, participantsupps1, by = "Participant")
merged_ph_brmps <- merged_ph_brmps %>%
  select(starts_with("Participant"), starts_with("ph.wk")) %>%
  mutate(ph.diff = ph.wk3 - ph.wk1) %>%
  arrange(by_group = ph.diff)

#Summary Statistics and Visualizations for BRMPS 2x Daily pH Change
nrow(merged_ph_brmps) 
t.test(as.vector(merged_ph_brmps$ph.wk1), 
       as.vector(merged_ph_brmps$ph.wk3),
       alternative = "greater",
       paired = TRUE,
       conf.level = 0.95)
summary(merged_ph_brmps)
ph_brmps1x_barplot <- barplot(merged_ph_brmps$ph.diff, ylab = "Participants", 
        xlab = "pH (Week 3 - Week 1)", 
        main = "BRMPS 2x Daily pH Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(merged_ph_brmps)
```

#-------------------------------------------------------------------------------

#BRMPS 1x Daily Cohort 
##Acetate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
    filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "BRMPS 1xdaily 40") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_acetate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_acetate,participantsupps1,by="Participant") %>%
  rename(acetate.wk1 = Acetate_mmol_kg) %>%
  select(starts_with("Participant"),
        starts_with("acetate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_acetate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate")) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_acetate,participantsupps1,by="Participant") %>%
  rename(acetate.wk3 = Acetate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("acetate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
acetate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
acetate_brmps_diff <- acetate_brmps %>%
  mutate(acetate_diff = acetate.wk3 - acetate.wk1) %>%
  arrange(by = acetate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
acetate_wk1_vec <- as.vector(acetate_brmps$acetate.wk1)
acetate_wk3_vec <- as.vector(acetate_brmps$acetate.wk3)
paired_acetate_df <- data.frame(Week1 = acetate_wk1_vec, Week3 = acetate_wk3_vec)
acetatepairplot <- ggpaired(paired_acetate_df, cond1 = "Week1", cond2 = "Week3")
acetatepairplot$layers <- acetatepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(acetate_brmps$acetate.wk1), 
       as.vector(acetate_brmps$acetate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(acetate_brmps_diff)
acetatepairplot
acebrmps1x_barplot <- barplot(acetate_brmps_diff$acetate_diff, ylab = "Participants", xlab = "Acet. Conc. Change (Week 3 - Week 1)", 
        main = "BRMPS 1x Daily Acetate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(acetate_brmps_diff)
```

##Acetate Proportion in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
    filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "BRMPS 1xdaily 40") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_acetateprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week1") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(acetate_prop1 = Acetate_mmol_kg / scfa_sum) 

fecal1_brmpsprop <- merge(fecal1_acetateprop,participantsupps1,by="Participant") %>%
  rename(acetate_prop_wk1 = acetate_prop1) %>%
  select(starts_with("Participant"),
         starts_with("acetate_prop_wk1"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_acetateprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week3") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(acetate_prop3 = Acetate_mmol_kg / scfa_sum) 

fecal3_brmpsprop <- merge(fecal3_acetateprop,participantsupps1,by="Participant") %>%
  rename(acetate_prop_wk3 = acetate_prop3) %>%
  select(starts_with("Participant"),
         starts_with("acetate_prop_wk3"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
acetate_brmpsprop <- merge(fecal1_brmpsprop,fecal3_brmpsprop,
                           by="Participant") %>%
                     select(-ends_with("x"), -ends_with("y"))

#Appending Differences column to acetate-BRMPS df in form: week3-week1
acetate_brmps_diffprop <- acetate_brmpsprop %>%
  mutate(acetate_prop_diff = acetate_prop_wk3 - acetate_prop_wk1) %>%
  arrange(by = acetate_prop_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
acetate_wk1_vecprop <- as.vector(acetate_brmps_diffprop$acetate_prop_wk1)
acetate_wk3_vecprop <- as.vector(acetate_brmps_diffprop$acetate_prop_wk3)
paired_acetate_dfprop <- data.frame(Week1 = acetate_wk1_vecprop, 
                                    Week3 = acetate_wk3_vecprop)
acetatepairplotprop <- ggpaired(paired_acetate_dfprop, cond1 = "Week1", 
                                cond2 = "Week3")
acetatepairplotprop$layers <- acetatepairplotprop$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(acetate_brmpsprop$acetate_prop_wk1), 
       as.vector(acetate_brmpsprop$acetate_prop_wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(acetate_brmps_diffprop) # p-value = 0.0971. 
acetatepairplotprop
aceprop1x_barplot <- barplot(acetate_brmps_diffprop$acetate_prop_diff, ylab = "Participants", xlab = "Acetate Prop. Change (Week 3 - Week 1)", 
        main = "BRMPS 1x Daily Ace. Prop. Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(acetate_brmps_diffprop)
```

##Butyrate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
fullfecalsample <- read.csv("./fecalsample_a.csv")

#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
    filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "BRMPS 1xdaily 40") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_butyrate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Butyrate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_butyrate,participantsupps1,by="Participant") %>%
  rename(butyrate.wk1 = Butyrate_mmol_kg) %>%
  select(starts_with("Participant"),
        starts_with("butyrate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_butyrate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Butyrate")) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_butyrate,participantsupps1,by="Participant") %>%
  rename(butyrate.wk3 = Butyrate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("butyrate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
butyrate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
butyrate_brmps_diff <- butyrate_brmps %>%
  mutate(butyrate_diff = butyrate.wk3 - butyrate.wk1) %>%
  arrange(by = butyrate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
butyrate_wk1_vec <- as.vector(butyrate_brmps$butyrate.wk1)
butyrate_wk3_vec <- as.vector(butyrate_brmps$butyrate.wk3)
paired_butyrate_df <- data.frame(Week1 = butyrate_wk1_vec, Week3 = butyrate_wk3_vec)
butyratepairplot <- ggpaired(paired_butyrate_df, cond1 = "Week1", cond2 = "Week3")
butyratepairplot$layers <- butyratepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(butyrate_brmps$butyrate.wk1), 
       as.vector(butyrate_brmps$butyrate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(butyrate_brmps_diff)
butyratepairplot
butbrmps1x_barplot <- barplot(butyrate_brmps_diff$butyrate_diff, ylab = "Participants", xlab = "Butyr. Conc. Change (Week 3 - Week 1)", 
        main = "BRMPS 1x Daily Butyrate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(butyrate_brmps_diff)
```

##Propionate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
    filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "BRMPS 1xdaily 40") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_propionate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Propionate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_propionate,participantsupps1,by="Participant") %>%
  rename(propionate.wk1 = Propionate_mmol_kg) %>%
  select(starts_with("Participant"),
        starts_with("propionate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_propionate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Propionate")) %>%
  filter(Week == "week3") %>%
  na.omit()

fecal3_brmps <- merge(fecal3_propionate,participantsupps1,by="Participant") %>%
  rename(propionate.wk3 = Propionate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("propionate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
propionate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
propionate_brmps_diff <- propionate_brmps %>%
  mutate(propionate_diff = propionate.wk3 - propionate.wk1) %>%
  arrange(by = propionate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
propionate_wk1_vec <- as.vector(propionate_brmps$propionate.wk1)
propionate_wk3_vec <- as.vector(propionate_brmps$propionate.wk3)
paired_propionate_df <- data.frame(Week1 = propionate_wk1_vec, Week3 = propionate_wk3_vec)
propionatepairplot <- ggpaired(paired_propionate_df, cond1 = "Week1", cond2 = "Week3")
propionatepairplot$layers <- propionatepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(propionate_brmps$propionate.wk1), 
       as.vector(propionate_brmps$propionate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(propionate_brmps_diff)
propionatepairplot
probrmps1x_barplot <- barplot(propionate_brmps_diff$propionate_diff, ylab = "Participants", xlab = "Propio. Conc. Change (Week 3 - Week 1)", 
        main = "BRMPS 1x Daily Propionate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(propionate_brmps_diff)
```

##Total SCFA Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
    filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "BRMPS 1xdaily 40") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_total <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate")) %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_total,participantsupps1,by="Participant") %>%
  rename(total.wk1 = scfa_sum) %>%
  select(starts_with("Participant"),
        starts_with("total"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_total <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate")) %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_total,participantsupps1,by="Participant") %>%
  rename(total.wk3 = scfa_sum) %>%
  select(starts_with("Participant"),
         starts_with("total"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
total_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
total_brmps_diff <- total_brmps %>%
  mutate(total_diff = total.wk3 - total.wk1) %>%
  arrange(by = total_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
total_wk1_vec <- as.vector(total_brmps$total.wk1)
total_wk3_vec <- as.vector(total_brmps$total.wk3)
paired_total_df <- data.frame(Week1 = total_wk1_vec, Week3 = total_wk3_vec)
totalpairplot <- ggpaired(paired_total_df, cond1 = "Week1", cond2 = "Week3")
totalpairplot$layers <- totalpairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(total_brmps$total.wk1), 
       as.vector(total_brmps$total.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(total_brmps_diff)
totalpairplot
totbrmps1x_barplot <- barplot(total_brmps_diff$total_diff, ylab = "Participants", xlab = "Total Conc. Change (Week 3 - Week 1)", 
        main = "BRMPS 1x Daily Total Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(total_brmps_diff)
```

##pH Change in BRMPS 1x Cohort from Week 1 to Week 3
```{r}
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015") %>%
  rename(Participant = Name) %>%
  filter(Supplement.consumed == "BRMPS 1xdaily 40")

#Making Week 1 BRMPS 1x-pH Data Frame 
ph_brmps_wk1 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week1") %>%
  na.omit() %>%
  rename(ph.wk1 = pH)

#Making Week 3 BRMPS 1x-pH Data Frame 
ph_brmps_wk3 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week3") %>%
  na.omit() %>%
  rename(ph.wk3 = pH)

#Combining Week 1 and Week 3 BRMPS 1x Daily Cohort 
merged_ph_wks <- merge(ph_brmps_wk1, ph_brmps_wk3, by = "Participant") 
merged_ph_brmps<- merge(merged_ph_wks, participantsupps1, by = "Participant")
merged_ph_brmps <- merged_ph_brmps %>%
  select(starts_with("Participant"), starts_with("ph.wk")) %>%
  mutate(ph.diff = ph.wk3 - ph.wk1) %>%
  arrange(by_group = ph.diff)

#Summary Statistics and Visualizations for BRMPS 1x Daily pH Change
nrow(merged_ph_brmps) 
t.test(as.vector(merged_ph_brmps$ph.wk1), 
       as.vector(merged_ph_brmps$ph.wk3),
       alternative = "greater",
       paired = TRUE,
       conf.level = 0.95)
summary(merged_ph_brmps)
ph_brmps1x_barplot <- barplot(merged_ph_brmps$ph.diff, ylab = "Participants", 
        xlab = "pH (Week 3 - Week 1)", 
        main = "BRMPS 1x Daily pH Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(merged_ph_brmps)
```

#-------------------------------------------------------------------------------

#No Supplement Cohort 
##Acetate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily None Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Supplement.consumed == "none") %>%
  rename(Participant = Name)


#Making Week 1 Data Frame from None and Acetate
fecal1_acetate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_none <- merge(fecal1_acetate,participantsupps1,by="Participant") %>%
  rename(acetate.wk1 = Acetate_mmol_kg) %>%
  select(starts_with("Participant"),
        starts_with("acetate"))

#Making Week 3 Data Frame from None and Acetate
fecal3_acetate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate")) %>%
  filter(Week == "week3") %>%
  na.omit()

fecal3_none <- merge(fecal3_acetate,participantsupps1,by="Participant") %>%
  rename(acetate.wk3 = Acetate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("acetate"))

#Combining Week 1 and Week 3 None-Acetate DFs for Larger DF
acetate_none <- merge(fecal1_none,fecal3_none,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-None df in form: week3-week1
acetate_none_diff <- acetate_none %>%
  mutate(acetate_diff = acetate.wk3 - acetate.wk1) %>%
  arrange(by = acetate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 None-Acetate Data
acetate_wk1_vec <- as.vector(acetate_none$acetate.wk1)
acetate_wk3_vec <- as.vector(acetate_none$acetate.wk3)
paired_acetate_df <- data.frame(Week1 = acetate_wk1_vec, Week3 = acetate_wk3_vec)
acetatepairplot <- ggpaired(paired_acetate_df, cond1 = "Week1", cond2 = "Week3")
acetatepairplot$layers <- acetatepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 None-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(acetate_none$acetate.wk1), 
       as.vector(acetate_none$acetate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(acetate_none_diff)
acetatepairplot
acebrmps1x_barplot <- barplot(acetate_none_diff$acetate_diff, ylab = "Participants", xlab = "Acet. Conc. Change (Week 3 - Week 1)", 
        main = "No Supp. Daily Acetate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(acetate_none_diff)
```

##Butyrate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
fullfecalsample <- read.csv("./fecalsample_a.csv")

#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily None Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Supplement.consumed == "none") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from None and Acetate
fecal1_butyrate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Butyrate")) %>%
  filter(Week == "week1") %>%
  na.omit()

fecal1_none <- merge(fecal1_butyrate,participantsupps1,by="Participant") %>%
  rename(butyrate.wk1 = Butyrate_mmol_kg) %>%
  select(starts_with("Participant"),
        starts_with("butyrate"))

#Making Week 3 Data Frame from None and Acetate
fecal3_butyrate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Butyrate")) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_none <- merge(fecal3_butyrate,participantsupps1,by="Participant") %>%
  rename(butyrate.wk3 = Butyrate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("butyrate"))

#Combining Week 1 and Week 3 None-Acetate DFs for Larger DF
butyrate_none <- merge(fecal1_none,fecal3_none,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-None df in form: week3-week1
butyrate_none_diff <- butyrate_none %>%
  mutate(butyrate_diff = butyrate.wk3 - butyrate.wk1) %>%
  arrange(by = butyrate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 None-Acetate Data
butyrate_wk1_vec <- as.vector(butyrate_none$butyrate.wk1)
butyrate_wk3_vec <- as.vector(butyrate_none$butyrate.wk3)
paired_butyrate_df <- data.frame(Week1 = butyrate_wk1_vec, Week3 = butyrate_wk3_vec)
butyratepairplot <- ggpaired(paired_butyrate_df, cond1 = "Week1", cond2 = "Week3")
butyratepairplot$layers <- butyratepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 None-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(butyrate_none$butyrate.wk1), 
       as.vector(butyrate_none$butyrate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(butyrate_none_diff)
butyratepairplot
butbrmps1x_barplot <- barplot(butyrate_none_diff$butyrate_diff, ylab = "Participants", xlab = "But. Conc. Change (Week 3 - Week 1)", 
        main = "No Supp. Daily Butyrate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(butyrate_none_diff)
```

##Propionate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily None Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Supplement.consumed == "none") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from None and Acetate
fecal1_propionate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Propionate")) %>%
  filter(Week == "week1") %>%
  na.omit()

fecal1_none <- merge(fecal1_propionate,participantsupps1,by="Participant") %>%
  rename(propionate.wk1 = Propionate_mmol_kg) %>%
  select(starts_with("Participant"),
        starts_with("propionate"))

#Making Week 3 Data Frame from None and Acetate
fecal3_propionate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Propionate")) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_none <- merge(fecal3_propionate,participantsupps1,by="Participant") %>%
  rename(propionate.wk3 = Propionate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("propionate"))

#Combining Week 1 and Week 3 None-Acetate DFs for Larger DF
propionate_none <- merge(fecal1_none,fecal3_none,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-None df in form: week3-week1
propionate_none_diff <- propionate_none %>%
  mutate(propionate_diff = propionate.wk3 - propionate.wk1) %>%
  arrange(by = propionate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 None-Acetate Data
propionate_wk1_vec <- as.vector(propionate_none$propionate.wk1)
propionate_wk3_vec <- as.vector(propionate_none$propionate.wk3)
paired_propionate_df <- data.frame(Week1 = propionate_wk1_vec, Week3 = propionate_wk3_vec)
propionatepairplot <- ggpaired(paired_propionate_df, cond1 = "Week1", cond2 = "Week3")
propionatepairplot$layers <- propionatepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 None-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(propionate_none$propionate.wk1), 
       as.vector(propionate_none$propionate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(propionate_none_diff)
propionatepairplot
probrmps1x_barplot <- barplot(propionate_none_diff$propionate_diff, ylab = "Participants", xlab = "Pro. Conc. Change (Week 3 - Week 1)", 
        main = "No Supp. Daily Propionate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(propionate_none_diff)
```

##Total SCFA Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and None Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Supplement.consumed == "none") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from None and Acetate
fecal1_total <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate")) %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_none <- merge(fecal1_total,participantsupps1,by="Participant") %>%
  rename(total.wk1 = scfa_sum) %>%
  select(starts_with("Participant"),
         starts_with("total"))

#Making Week 3 Data Frame from None and Acetate
fecal3_total <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate")) %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_none <- merge(fecal3_total,participantsupps1,by="Participant") %>%
  rename(total.wk3 = scfa_sum) %>%
  select(starts_with("Participant"),
         starts_with("total"))

#Combining Week 1 and Week 3 None-Acetate DFs for Larger DF
total_none <- merge(fecal1_none,fecal3_none,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-None df in form: week3-week1
total_none_diff <- total_none %>%
  mutate(total_diff = total.wk3 - total.wk1) %>%
  arrange(by = total_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 None-Acetate Data
total_wk1_vec <- as.vector(total_none$total.wk1)
total_wk3_vec <- as.vector(total_none$total.wk3)
paired_total_df <- data.frame(Week1 = total_wk1_vec, Week3 = total_wk3_vec)
totalpairplot <- ggpaired(paired_total_df, cond1 = "Week1", cond2 = "Week3")
totalpairplot$layers <- totalpairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 None-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(total_none$total.wk1), 
       as.vector(total_none$total.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(total_none_diff)
totalpairplot
totnonw_barplot <- barplot(total_none_diff$total_diff, ylab = "Participants", xlab = "Total Conc. Change (Week 3 - Week 1)", 
        main = "No Supp. Total Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(total_none_diff)
```

##pH Change in No-Supp. Cohort from Week 1 to Week 3
```{r}
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015") %>%
  rename(Participant = Name) %>%
  filter(Supplement.consumed == "none")

#Making Week 1 BRMPS 1x-pH Data Frame 
ph_none_wk1 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week1") %>%
  na.omit() %>%
  rename(ph.wk1 = pH)

#Making Week 3 BRMPS 1x-pH Data Frame 
ph_none_wk3 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week3") %>%
  na.omit() %>%
  rename(ph.wk3 = pH)

#Combining Week 1 and Week 3 BRMPS 1x Daily Cohort 
merged_ph_wks <- merge(ph_none_wk1, ph_none_wk3, by = "Participant") 
merged_ph_none<- merge(merged_ph_wks, participantsupps1, by = "Participant")
merged_ph_none <- merged_ph_none %>%
  select(starts_with("Participant"), starts_with("ph.wk")) %>%
  mutate(ph.diff = ph.wk3 - ph.wk1) %>%
  arrange(by_group = ph.diff)

#Summary Statistics and Visualizations for BRMPS 1x Daily pH Change
nrow(merged_ph_none) 
t.test(as.vector(merged_ph_none$ph.wk1), 
       as.vector(merged_ph_none$ph.wk3),
       alternative = "greater",
       paired = TRUE,
       conf.level = 0.95)
summary(merged_ph_none)
ph_none_barplot <- barplot(merged_ph_none$ph.diff, ylab = "Participants", 
        xlab = "pH (Week 3 - Week 1)", 
        main = "None pH Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(merged_ph_none)
```

#-------------------------------------------------------------------------------
#Bristol vs. PH
##Bristol vs. PH for Week 1 and Week 3 (Participants Averaged)
```{r}
#Reading in Fecal Sample Data
fullfecalsample <- read.csv("./fecalsample_a.csv")

fecal_clean <- fullfecalsample %>%
  na.omit() %>%
  select(starts_with("Participant"), 
         starts_with("Week"),
         starts_with("pH"),
         starts_with("Bristol")) %>%
  filter(Week == "week1" | Week == "week3") 

#Preparing Data for Scatterplot Creation 
fecal_clean$pH = as.numeric(fecal_clean$pH)
fecal_clean$Bristol = as.numeric(fecal_clean$Bristol)

#Creating Scatterplot for SCFA vs. PH
ggplot(fecal_clean, aes(x=pH, y=Bristol)) +
  geom_point(alpha=0.3) +
  ggtitle("PH vs. BRISTOL for All Weeks") +
  labs(y= "BRISTOL", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH without Points
ggplot(fecal_clean, aes(x=pH, y=Bristol)) +
  geom_smooth() +
  ggtitle("BRISTOL vs. PH for All Weeks") +
  labs(y= "BRISTOL", x = "Average PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH with Points
ggplot(fecal_clean, aes(x=pH, y=Bristol)) +
  geom_point() +
  geom_smooth() +
  ggtitle("BRISTOL vs. PH for All Weeks") +
  labs(y= "BRISTOL", x = "Average PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Correlation Test
cor(as.vector(fecal_clean$pH), as.vector(fecal_clean$Bristol),
    method = "pearson") #Returns -0.01719789. Low negative correlation 

#Significance Statistics
brisph_norm<- qqplot(as.vector(fecal_clean$pH), as.vector(fecal_clean$Bristol))
bartlett.test(Bristol ~ Week, data = fecal_clean)
cor.test(as.vector(fecal_clean$pH), as.vector(fecal_clean$Bristol), 
         alternative = "less", method = 'spearman')
```

##Bristol vs. PH for Week 1 and Week 3 (Participants Unaveraged)
```{r}
#Reading in Fecal Sample Data
fullfecalsample_unaveraged <- read.csv("./fecalsample_ua.csv")
fecal_clean_bristol <- fullfecalsample_unaveraged %>%
  na.omit() %>%
  select(starts_with("Participant"), 
         starts_with("Week"),
         starts_with("pH"),
         starts_with("Bristol")) %>%
  filter(Week == "week1" | Week == "week3") 

#Preparing Data for Scatterplot Creation 
fecal_clean_bristol$pH = as.numeric(fecal_clean_bristol$pH)
fecal_clean_bristol$Bristol = as.numeric(fecal_clean_bristol$Bristol)

#Creating Scatterplot for SCFA vs. PH
ggplot(fecal_clean_bristol, aes(x=pH, y=Bristol)) +
  geom_point(alpha=0.3) +
  ggtitle("PH vs. Bristol for All Weeks") +
  labs(y= "Bristol", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH without Points
ggplot(fecal_clean_bristol, aes(x=pH, y=Bristol)) +
  geom_smooth() +
  ggtitle("Bristol vs. PH for All Weeks") +
  labs(y= "Bristol", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH with Points
ggplot(fecal_clean_bristol, aes(x=pH, y=Bristol)) +
  geom_point() +
  geom_smooth() +
  ggtitle("Bristol vs. PH for All Weeks") +
  labs(y= "Bristol", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Correlation Test
cor(as.vector(fecal_clean_bristol$pH), as.vector(fecal_clean_bristol$Bristol),
    method = "spearman") 

#Significance Statistics
qqplot(as.vector(fecal_clean_bristol$pH), as.vector(fecal_clean_bristol$Bristol))
bartlett.test(Bristol ~ Week, data = fecal_clean_bristol)
cor.test(as.vector(fecal_clean_bristol$pH),as.vector(fecal_clean_bristol$Bristol), alternative = "less", method = 'spearman')
```

##Bristol vs. PH for Week 1 Only 
```{r}
#Reading in Fecal Sample Data
fullfecalsample <- read.csv("./fecalsample_a.csv")

fecal_clean <- fullfecalsample %>%
  na.omit() %>%
  select(starts_with("Participant"), 
         starts_with("Week"),
         starts_with("pH"),
         starts_with("Bristol")) %>%
  filter(Week == "week1") 

#Preparing Data for Scatterplot Creation 
fecal_clean$pH = as.numeric(fecal_clean$pH)
fecal_clean$Bristol = as.numeric(fecal_clean$Bristol)

#Creating Scatterplot for SCFA vs. PH
ggplot(fecal_clean, aes(x=pH, y=Bristol)) +
  geom_point(alpha=0.3) +
  ggtitle("PH vs. BRISTOL for All Weeks") +
  labs(y= "BRISTOL", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH without Points
ggplot(fecal_clean, aes(x=pH, y=Bristol)) +
  geom_smooth() +
  ggtitle("BRISTOL vs. PH for All Weeks") +
  labs(y= "BRISTOL", x = "Average PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH with Points
ggplot(fecal_clean, aes(x=pH, y=Bristol)) +
  geom_point() +
  geom_smooth() +
  ggtitle("BRISTOL vs. PH for All Weeks") +
  labs(y= "BRISTOL", x = "Average PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Correlation Test
cor(as.vector(fecal_clean$pH), as.vector(fecal_clean$Bristol),
    method = "pearson") #Returns -0.01719789. Low negative correlation 

#Significance Statistics
brisph_norm<- qqplot(as.vector(fecal_clean$pH), as.vector(fecal_clean$Bristol))
cor.test(as.vector(fecal_clean$pH), as.vector(fecal_clean$Bristol), 
         alternative = "less", method = 'spearman')
```

##Bristol vs. PH for Week 1 Only (Participants Unaveraged)
```{r}
#Reading in Fecal Sample Data
fullfecalsample_unaveraged <- read.csv("./fecalsample_ua.csv")
fecal_clean_bristol <- fullfecalsample_unaveraged %>%
  na.omit() %>%
  select(starts_with("Participant"), 
         starts_with("Week"),
         starts_with("pH"),
         starts_with("Bristol")) %>%
  filter(Week == "week1") 

#Preparing Data for Scatterplot Creation 
fecal_clean_bristol$pH = as.numeric(fecal_clean_bristol$pH)
fecal_clean_bristol$Bristol = as.numeric(fecal_clean_bristol$Bristol)

#Creating Scatterplot for SCFA vs. PH
ggplot(fecal_clean_bristol, aes(x=pH, y=Bristol)) +
  geom_point(alpha=0.3) +
  ggtitle("PH vs. Bristol for All Weeks") +
  labs(y= "Bristol", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH without Points
ggplot(fecal_clean_bristol, aes(x=pH, y=Bristol)) +
  geom_smooth() +
  ggtitle("Bristol vs. PH for All Weeks") +
  labs(y= "Bristol", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH with Points
ggplot(fecal_clean_bristol, aes(x=pH, y=Bristol)) +
  geom_point() +
  geom_smooth() +
  ggtitle("Bristol vs. PH for All Weeks") +
  labs(y= "Bristol", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Correlation Test
cor(as.vector(fecal_clean_bristol$pH), as.vector(fecal_clean_bristol$Bristol),
    method = "spearman") 

#Significance Statistics
qqplot(as.vector(fecal_clean_bristol$pH), as.vector(fecal_clean_bristol$Bristol))
cor.test(as.vector(fecal_clean_bristol$pH),as.vector(fecal_clean_bristol$Bristol), alternative = "less", method = 'spearman')
```

##Bristol vs. PH for Week 3 Only 
```{r}
#Reading in Fecal Sample Data
fullfecalsample <- read.csv("./fecalsample_a.csv")

fecal_clean <- fullfecalsample %>%
  na.omit() %>%
  select(starts_with("Participant"), 
         starts_with("Week"),
         starts_with("pH"),
         starts_with("Bristol")) %>%
  filter(Week == "week3") 

#Preparing Data for Scatterplot Creation 
fecal_clean$pH = as.numeric(fecal_clean$pH)
fecal_clean$Bristol = as.numeric(fecal_clean$Bristol)

#Creating Scatterplot for SCFA vs. PH
ggplot(fecal_clean, aes(x=pH, y=Bristol)) +
  geom_point(alpha=0.3) +
  ggtitle("PH vs. BRISTOL for All Weeks") +
  labs(y= "BRISTOL", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH without Points
ggplot(fecal_clean, aes(x=pH, y=Bristol)) +
  geom_smooth() +
  ggtitle("BRISTOL vs. PH for All Weeks") +
  labs(y= "BRISTOL", x = "Average PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH with Points
ggplot(fecal_clean, aes(x=pH, y=Bristol)) +
  geom_point() +
  geom_smooth() +
  ggtitle("BRISTOL vs. PH for All Weeks") +
  labs(y= "BRISTOL", x = "Average PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Correlation Test
cor(as.vector(fecal_clean$pH), as.vector(fecal_clean$Bristol),
    method = "spearman") #Returns -0.01719789. Low negative correlation 

#Significance Statistics
brisph_norm<- qqplot(as.vector(fecal_clean$pH), as.vector(fecal_clean$Bristol))
cor.test(as.vector(fecal_clean$pH), as.vector(fecal_clean$Bristol), 
         alternative = "less", method = 'pearson')
```

##Bristol vs. PH for Week 3 Only (Participants Unaveraged)
```{r}
#Reading in Fecal Sample Data
fullfecalsample_unaveraged <- read.csv("./fecalsample_ua.csv")
fecal_clean_bristol <- fullfecalsample_unaveraged %>%
  na.omit() %>%
  select(starts_with("Participant"), 
         starts_with("Week"),
         starts_with("pH"),
         starts_with("Bristol")) %>%
  filter(Week == "week3") 

#Preparing Data for Scatterplot Creation 
fecal_clean_bristol$pH = as.numeric(fecal_clean_bristol$pH)
fecal_clean_bristol$Bristol = as.numeric(fecal_clean_bristol$Bristol)

#Creating Scatterplot for SCFA vs. PH
ggplot(fecal_clean_bristol, aes(x=pH, y=Bristol)) +
  geom_point(alpha=0.3) +
  ggtitle("PH vs. Bristol for All Weeks") +
  labs(y= "Bristol", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH without Points
ggplot(fecal_clean_bristol, aes(x=pH, y=Bristol)) +
  geom_smooth() +
  ggtitle("Bristol vs. PH for All Weeks") +
  labs(y= "Bristol", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Creating Smoothed Conditional Means for SCFA vs. PH with Points
ggplot(fecal_clean_bristol, aes(x=pH, y=Bristol)) +
  geom_point() +
  geom_smooth() +
  ggtitle("Bristol vs. PH for All Weeks") +
  labs(y= "Bristol", x = "PH") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) 

#Correlation Test
cor(as.vector(fecal_clean_bristol$pH), as.vector(fecal_clean_bristol$Bristol),
    method = "spearman") 

#Significance Statistics
qqplot(as.vector(fecal_clean_bristol$pH), as.vector(fecal_clean_bristol$Bristol))
cor.test(as.vector(fecal_clean_bristol$pH),as.vector(fecal_clean_bristol$Bristol), alternative = "less", method = 'spearman')
```

#-------------------------------------------------------------------------------

#HiMaize 2x Daily Cohort 
##Acetate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily HiMaize Participants
participantsupps <- participantsupps %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "HiMaize 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from HiMaize and Acetate
fecal1_acetate <- read.csv("./fecalsample_a.csv") %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate_mmol_kg")) %>%
  filter(Week == "week1") %>%
  na.omit()

fecal1_himaize <- merge(fecal1_acetate,participantsupps,by="Participant") %>%
  rename(acetate.wk1 = Acetate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("acetate.wk1"))

#Making Week 3 Data Frame from HiMaize and Acetate
fecal3_acetate <- read.csv("./fecalsample_a.csv") %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate_mmol_kg")) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_himaize <- merge(fecal3_acetate,participantsupps,by="Participant") %>%
  rename(acetate.wk3 = Acetate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("acetate.wk3"))

#Combining Week 1 and Week 3 HiMaize-Acetate DFs for Larger DF
acetate_himaize <- merge(fecal1_himaize,fecal3_himaize,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-HiMaize df in form: week3-week1
acetate_himaize_diff <- acetate_himaize %>%
  mutate(acetate_diff = acetate.wk3 - acetate.wk1) %>%
  arrange(by = acetate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 HiMaize-Acetate Data
acetate_wk1_vec <- as.vector(acetate_himaize$acetate.wk1)
acetate_wk3_vec <- as.vector(acetate_himaize$acetate.wk3)
paired_acetate_df <- data.frame(Week1 = acetate_wk1_vec, Week3 = acetate_wk3_vec)
acetatepairplot <- ggpaired(paired_acetate_df, cond1 = "Week1", cond2 = "Week3")
acetatepairplot$layers <- acetatepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 HiMaize-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(acetate_himaize$acetate.wk1), 
       as.vector(acetate_himaize$acetate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(acetate_himaize_diff)
acetatepairplot
acebrmps2x_barplot <- barplot(acetate_himaize_diff$acetate_diff, ylab = "Participants", xlab = "Acetate. Conc. Change (Week 3 - Week 1)", 
        main = "HiMaize 2x Daily Acetate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(acetate_himaize_diff)
```

##Acetate Proportion in Week 1 vs. Week 3
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "HiMaize 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_acetateprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week1") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(acetate_prop1 = Acetate_mmol_kg / scfa_sum) 

fecal1_himaizeprop <- merge(fecal1_acetateprop,participantsupps1,by="Participant") %>%
  rename(acetate_prop_wk1 = acetate_prop1) %>%
  select(starts_with("Participant"),
         starts_with("acetate_prop_wk1"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_acetateprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week3") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(acetate_prop3 = Acetate_mmol_kg / scfa_sum) 

fecal3_himaizeprop <- merge(fecal3_acetateprop,participantsupps1,by="Participant") %>%
  rename(acetate_prop_wk3 = acetate_prop3) %>%
  select(starts_with("Participant"),
         starts_with("acetate_prop_wk3"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
acetate_himaizeprop <- merge(fecal1_himaizeprop,fecal3_himaizeprop,
                           by="Participant") %>%
                     select(-ends_with("x"), -ends_with("y"))

#Appending Differences column to acetate-BRMPS df in form: week3-week1
acetate_himaize_diffprop <- acetate_himaizeprop %>%
  mutate(acetate_prop_diff = acetate_prop_wk3 - acetate_prop_wk1) %>%
  arrange(by = acetate_prop_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
acetate_wk1_vecprop <- as.vector(acetate_himaize_diffprop$acetate_prop_wk1)
acetate_wk3_vecprop <- as.vector(acetate_himaize_diffprop$acetate_prop_wk3)
paired_acetate_dfprop <- data.frame(Week1 = acetate_wk1_vecprop, 
                                    Week3 = acetate_wk3_vecprop)
acetatepairplotprop <- ggpaired(paired_acetate_dfprop, cond1 = "Week1", 
                                cond2 = "Week3")
acetatepairplotprop$layers <- acetatepairplotprop$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(acetate_himaizeprop$acetate_prop_wk1), 
       as.vector(acetate_himaizeprop$acetate_prop_wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(acetate_himaize_diffprop) # p-value = 0.0971. 
acetatepairplotprop
aceprop1x_barplot <- barplot(acetate_himaize_diffprop$acetate_prop_diff, ylab = "Participants", xlab = "Acetate Prop. Change (Week 3 - Week 1)", 
        main = "HiMaize 2x Daily Ace. Prop. Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(acetate_himaize_diffprop)
```

##Butyrate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
fullfecalsample <- read.csv("./fecalsample_a.csv")

#Making and Cleaning data frame with Participants by Supplement
participantsupps <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily HiMaize Participants
participantsupps <- participantsupps %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "HiMaize 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from HiMaize and Acetate
fecal1_butyrate <- read.csv("./fecalsample_a.csv") %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Butyrate_mmol_kg")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_himaize <- merge(fecal1_butyrate,participantsupps,by="Participant") %>%
  rename(butyrate.wk1 = Butyrate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("butyrate.wk1"))

#Making Week 3 Data Frame from HiMaize and Acetate
fecal3_butyrate <- read.csv("./fecalsample_a.csv") %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Butyrate_mmol_kg")) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_himaize <- merge(fecal3_butyrate,participantsupps,by="Participant") %>%
  rename(butyrate.wk3 = Butyrate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("butyrate.wk3"))

#Combining Week 1 and Week 3 HiMaize-Acetate DFs for Larger DF
butyrate_himaize <- merge(fecal1_himaize,fecal3_himaize,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-HiMaize df in form: week3-week1
butyrate_himaize_diff <- butyrate_himaize %>%
  mutate(butyrate_diff = butyrate.wk3 - butyrate.wk1) %>%
  arrange(by = butyrate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 HiMaize-Acetate Data
butyrate_wk1_vec <- as.vector(butyrate_himaize$butyrate.wk1)
butyrate_wk3_vec <- as.vector(butyrate_himaize$butyrate.wk3)
paired_butyrate_df <- data.frame(Week1 = butyrate_wk1_vec, Week3 = butyrate_wk3_vec)
butyratepairplot <- ggpaired(paired_butyrate_df, cond1 = "Week1", cond2 = "Week3")
butyratepairplot$layers <- butyratepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 HiMaize-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(butyrate_himaize$butyrate.wk1), 
       as.vector(butyrate_himaize$butyrate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(butyrate_himaize_diff)
butyratepairplot
butbrmps2x_barplot <- barplot(butyrate_himaize_diff$butyrate_diff, ylab = "Participants", xlab = "Butyrate. Conc. Change (Week 3 - Week 1)", 
        main = "HiMaize 2x Daily Butyrate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(butyrate_himaize_diff)
```

##Propionate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily HiMaize Participants
participantsupps <- participantsupps %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "HiMaize 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from HiMaize and Acetate
fecal1_propionate <- read.csv("./fecalsample_a.csv") %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Propionate_mmol_kg")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_himaize <- merge(fecal1_propionate,participantsupps,by="Participant") %>%
  rename(propionate.wk1 = Propionate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("propionate.wk1"))

#Making Week 3 Data Frame from HiMaize and Acetate
fecal3_propionate <- read.csv("./fecalsample_a.csv") %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Propionate_mmol_kg")) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_himaize <- merge(fecal3_propionate,participantsupps,by="Participant") %>%
  rename(propionate.wk3 = Propionate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("propionate.wk3"))

#Combining Week 1 and Week 3 HiMaize-Acetate DFs for Larger DF
propionate_himaize <- merge(fecal1_himaize,fecal3_himaize,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-HiMaize df in form: week3-week1
propionate_himaize_diff <- propionate_himaize %>%
  mutate(propionate_diff = propionate.wk3 - propionate.wk1) %>%
  arrange(by = propionate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 HiMaize-Acetate Data
propionate_wk1_vec <- as.vector(propionate_himaize$propionate.wk1)
propionate_wk3_vec <- as.vector(propionate_himaize$propionate.wk3)
paired_propionate_df <- data.frame(Week1 = propionate_wk1_vec, Week3 = propionate_wk3_vec)
propionatepairplot <- ggpaired(paired_propionate_df, cond1 = "Week1", cond2 = "Week3")
propionatepairplot$layers <- propionatepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 HiMaize-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(propionate_himaize$propionate.wk1), 
       as.vector(propionate_himaize$propionate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(propionate_himaize_diff)
propionatepairplot
propbrmps2x_barplot <- barplot(propionate_himaize_diff$propionate_diff, ylab = "Participants", xlab = "Prop. Conc. Change (Week 3 - Week 1)", 
        main = "HiMaize 2x Daily Prop. Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(propionate_himaize_diff)
```

##Total SCFA Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps <- participantsupps %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "HiMaize 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Total SCFAs
fecal1_tot <- read.csv("./fecalsample_a.csv") %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate_mmol_kg"),
         starts_with("Butyrate_mmol_kg"),
         starts_with("Propionate_mmol_kg")) %>%
  filter(Week == "week1") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5]))

fecal1_himaize <- merge(fecal1_tot,participantsupps,by="Participant") %>%
  rename(scfa.wk1 = scfa_sum) %>%
  select(starts_with("Participant"),
         starts_with("scfa.wk1"))

#Making Week 3 Data Frame from BRMPS and Total SCFAs
fecal3_tot <- read.csv("./fecalsample_a.csv") %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate_mmol_kg"),
         starts_with("Butyrate_mmol_kg"),
         starts_with("Propionate_mmol_kg")) %>%
  filter(Week == "week3") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5]))

fecal3_himaize <- merge(fecal3_tot,participantsupps,by="Participant") %>%
  rename(scfa.wk3 = scfa_sum) %>%
  select(starts_with("Participant"),
         starts_with("scfa.wk3"))

#Combining Week 1 and Week 3 HiMaize-Acetate DFs for Larger DF
tot_himaize <- merge(fecal1_himaize,fecal3_himaize,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-HiMaize df in form: week3-week1
tot_himaize_diff <- tot_himaize %>%
  mutate(tot_diff = scfa.wk3 - scfa.wk1) %>%
  arrange(by = tot_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Total SCFA Data
#tot_wk1_vec <- as.vector(tot_himaize$scfa.wk1)
#tot_wk3_vec <- as.vector(tot_himaize$scfa.wk3)
#paired_tot_df <- data.frame(Week1 = tot_wk1_vec, Week3 = tot_wk3_vec)
#totpairplot <- ggpaired(paired_tot_df, cond1 = "Week1", cond2 = "Week3")
#totpairplot$layers <- totpairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Total SCFA
##t-test means (wk1-wk3) difference
t.test(as.vector(tot_himaize$scfa.wk1), 
       as.vector(tot_himaize$scfa.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(tot_himaize_diff)
#totpairplot
totscfahimaize2x_barplot <- barplot(tot_himaize_diff$tot_diff, ylab = "Participants",
      xlab = "Tot. SCFA Conc. Change (Wk 3 - Wk 1)", 
        main = "BRMPS 2x Daily Tot. SCFA Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(tot_himaize_diff)
```

```{r}
#Making Week 1 BRMPS 1x-pH Data Frame 
ph_himaize_wk1 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week1") %>%
  na.omit() %>%
  rename(ph.wk1 = pH)

#Making Week 3 BRMPS 1x-pH Data Frame 
ph_himaize_wk3 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week3") %>%
  na.omit() %>%
  rename(ph.wk3 = pH)

#Combining Week 1 and Week 3 BRMPS 1x Daily Cohort 
merged_ph_wks <- merge(ph_himaize_wk1, ph_himaize_wk3, by = "Participant") 
merged_ph_himaize<- merge(merged_ph_wks, participantsupps1, by = "Participant")
merged_ph_himaize <- merged_ph_himaize %>%
  select(starts_with("Participant"), starts_with("ph.wk")) %>%
  mutate(ph.diff = ph.wk3 - ph.wk1) %>%
  arrange(by_group = ph.diff)

#Summary Statistics and Visualizations for BRMPS 1x Daily pH Change
nrow(merged_ph_brmps) 
t.test(as.vector(merged_ph_himaize$ph.wk1), 
       as.vector(merged_ph_himaize$ph.wk3),
       alternative = "greater",
       paired = TRUE,
       conf.level = 0.95)
summary(merged_ph_brmps)
ph_brmps1x_barplot <- barplot(merged_ph_himaize$ph.diff, ylab = "Participants", 
        xlab = "pH (Week 3 - Week 1)", 
        main = "HiMaize 2x Daily pH Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(merged_ph_himaize)
```

##pH Change in HiMaize 2x Cohort from Week 1 to Week 3
```{r}
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015") %>%
  rename(Participant = Name) %>%
  filter(Supplement.consumed == "HiMaize 2xdaily 20")

#Making Week 1 BRMPS 1x-pH Data Frame 
ph_himaize_wk1 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week1") %>%
  na.omit() %>%
  rename(ph.wk1 = pH)

#Making Week 3 BRMPS 1x-pH Data Frame 
ph_himaize_wk3 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week3") %>%
  na.omit() %>%
  rename(ph.wk3 = pH)
View(ph_himaize_wk3)
View(participantsupps)

#Combining Week 1 and Week 3 BRMPS 1x Daily Cohort 
merged_ph_wks <- merge(ph_himaize_wk1, ph_himaize_wk3, by = "Participant") 
merged_ph_himaize<- merge(merged_ph_wks, participantsupps, by = "Participant")
merged_ph_himaize <- merged_ph_himaize %>%
  select(starts_with("Participant"), starts_with("ph.wk")) %>%
  mutate(ph.diff = ph.wk3 - ph.wk1) %>%
  arrange(by_group = ph.diff)

#Summary Statistics and Visualizations for BRMPS 1x Daily pH Change
nrow(merged_ph_himaize) 
t.test(as.vector(merged_ph_himaize$ph.wk1), 
       as.vector(merged_ph_himaize$ph.wk3),
       alternative = "greater",
       paired = TRUE,
       conf.level = 0.95)
summary(merged_ph_himaize)
ph_brmps1x_barplot <- barplot(merged_ph_himaize$ph.diff, ylab = "Participants", 
        xlab = "pH (Week 3 - Week 1)", 
        main = "HiMaize 2x Daily pH Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(merged_ph_brmps)
```
#-------------------------------------------------------------------------------
#LODAAT 1x Daily 20
##Acetate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
fullfecalsample <- read.csv("./fecalsample_a.csv")

#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 1xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_acetate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_acetate,participantsupps1,by="Participant") %>%
  rename(acetate.wk1 = Acetate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("acetate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_acetate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate")) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_acetate,participantsupps1,by="Participant") %>%
  rename(acetate.wk3 = Acetate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("acetate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
acetate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
acetate_brmps_diff <- acetate_brmps %>%
  mutate(acetate_diff = acetate.wk3 - acetate.wk1) %>%
  arrange(by = acetate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
acetate_wk1_vec <- as.vector(acetate_brmps$acetate.wk1)
acetate_wk3_vec <- as.vector(acetate_brmps$acetate.wk3)
paired_acetate_df <- data.frame(Week1 = acetate_wk1_vec, Week3 = acetate_wk3_vec)
acetatepairplot <- ggpaired(paired_acetate_df, cond1 = "Week1", cond2 = "Week3")
acetatepairplot$layers <- acetatepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(acetate_brmps$acetate.wk1), 
       as.vector(acetate_brmps$acetate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(acetate_brmps_diff)
acetatepairplot
acebrmps1x_barplot <- barplot(acetate_brmps_diff$acetate_diff, ylab = "Participants", xlab = "Acet. Conc. Change (Week 3 - Week 1)", 
        main = "LODAAT 1x Daily 20 Acetate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(acetate_brmps_diff)
```

##Acetate Proportion in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 1xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_acetateprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week1") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(acetate_prop1 = Acetate_mmol_kg / scfa_sum) 

fecal1_brmpsprop <- merge(fecal1_acetateprop,participantsupps1,by="Participant") %>%
  rename(acetate_prop_wk1 = acetate_prop1) %>%
  select(starts_with("Participant"),
         starts_with("acetate_prop_wk1"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_acetateprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week3") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(acetate_prop3 = Acetate_mmol_kg / scfa_sum) 

fecal3_brmpsprop <- merge(fecal3_acetateprop,participantsupps1,by="Participant") %>%
  rename(acetate_prop_wk3 = acetate_prop3) %>%
  select(starts_with("Participant"),
         starts_with("acetate_prop_wk3"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
acetate_brmpsprop <- merge(fecal1_brmpsprop,fecal3_brmpsprop,
                           by="Participant") %>%
                     select(-ends_with("x"), -ends_with("y"))

#Appending Differences column to acetate-BRMPS df in form: week3-week1
acetate_brmps_diffprop <- acetate_brmpsprop %>%
  mutate(acetate_prop_diff = acetate_prop_wk3 - acetate_prop_wk1) %>%
  arrange(by = acetate_prop_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
acetate_wk1_vecprop <- as.vector(acetate_brmps_diffprop$acetate_prop_wk1)
acetate_wk3_vecprop <- as.vector(acetate_brmps_diffprop$acetate_prop_wk3)
paired_acetate_dfprop <- data.frame(Week1 = acetate_wk1_vecprop, 
                                    Week3 = acetate_wk3_vecprop)
acetatepairplotprop <- ggpaired(paired_acetate_dfprop, cond1 = "Week1", 
                                cond2 = "Week3")
acetatepairplotprop$layers <- acetatepairplotprop$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(acetate_brmpsprop$acetate_prop_wk1), 
       as.vector(acetate_brmpsprop$acetate_prop_wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(acetate_brmps_diffprop) # p-value = 0.0971. 
acetatepairplotprop
aceprop1x_barplot <- barplot(acetate_brmps_diffprop$acetate_prop_diff, ylab = "Participants", xlab = "Acetate Prop. Change (Week 3 - Week 1)", 
        main = "LODAAT 1x Daily Ace. Prop. Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(acetate_brmps_diffprop)
```

##Butyrate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
fullfecalsample <- read.csv("./fecalsample_a.csv")

#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015") 

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 1xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_butyrate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Butyrate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_butyrate,participantsupps1,by="Participant") %>%
  rename(butyrate.wk1 = Butyrate_mmol_kg) %>%
  select(starts_with("Participant"),
        starts_with("butyrate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_butyrate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Butyrate")) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_butyrate,participantsupps1,by="Participant") %>%
  rename(butyrate.wk3 = Butyrate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("butyrate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
butyrate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round) 

#Appending Differences column to acetate-BRMPS df in form: week3-week1
butyrate_brmps_diff <- butyrate_brmps %>%
  mutate(butyrate_diff = butyrate.wk3 - butyrate.wk1) %>%
  arrange(by = butyrate_diff)

butyrate_brmps_diff <- butyrate_brmps_diff[-c(86, 87, 88), ]

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
butyrate_wk1_vec <- as.vector(butyrate_brmps$butyrate.wk1)
butyrate_wk3_vec <- as.vector(butyrate_brmps$butyrate.wk3)
paired_butyrate_df <- data.frame(Week1 = butyrate_wk1_vec, Week3 = butyrate_wk3_vec)
butyratepairplot <- ggpaired(paired_butyrate_df, cond1 = "Week1", cond2 = "Week3")
butyratepairplot$layers <- butyratepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(butyrate_brmps_diff$butyrate.wk1), 
       as.vector(butyrate_brmps_diff$butyrate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(butyrate_brmps_diff)
butyratepairplot
butbrmps1x_barplot <- barplot(butyrate_brmps_diff$butyrate_diff, ylab = "Participants", xlab = "Butyr. Conc. Change (Week 3 - Week 1)", 
        main = "LODAAT 1x Daily Butyrate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(butyrate_brmps_diff)
```

##Propionate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 1xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_propionate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Propionate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_propionate,participantsupps1,by="Participant") %>%
  rename(propionate.wk1 = Propionate_mmol_kg) %>%
  select(starts_with("Participant"),
        starts_with("propionate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_propionate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Propionate")) %>%
  filter(Week == "week3") %>%
  na.omit()

fecal3_brmps <- merge(fecal3_propionate,participantsupps1,by="Participant") %>%
  rename(propionate.wk3 = Propionate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("propionate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
propionate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
propionate_brmps_diff <- propionate_brmps %>%
  mutate(propionate_diff = propionate.wk3 - propionate.wk1) %>%
  arrange(by = propionate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
propionate_wk1_vec <- as.vector(propionate_brmps$propionate.wk1)
propionate_wk3_vec <- as.vector(propionate_brmps$propionate.wk3)
paired_propionate_df <- data.frame(Week1 = propionate_wk1_vec, Week3 = propionate_wk3_vec)
propionatepairplot <- ggpaired(paired_propionate_df, cond1 = "Week1", cond2 = "Week3")
propionatepairplot$layers <- propionatepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(propionate_brmps$propionate.wk1), 
       as.vector(propionate_brmps$propionate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(propionate_brmps_diff)
propionatepairplot
probrmps1x_barplot <- barplot(propionate_brmps_diff$propionate_diff, ylab = "Participants", xlab = "Propio. Conc. Change (Week 3 - Week 1)", 
        main = "LODAAT 1x Daily Propionate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(propionate_brmps_diff)
```
##Prop+But Proportion (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 1xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_bpprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week1") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(bp_prop1 = (Butyrate_mmol_kg+Propionate_mmol_kg) / scfa_sum) 
#View(fullfecalsample)

fecal1_brmpsprop <- merge(fecal1_bpprop,participantsupps1,by="Participant") %>%
  rename(bp_prop_wk1 = bp_prop1) %>%
  select(starts_with("Participant"),
         starts_with("bp_prop_wk1"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_bpprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week3") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(bp_prop3 = (Butyrate_mmol_kg+Propionate_mmol_kg) / scfa_sum) 

fecal3_brmpsprop <- merge(fecal3_bpprop,participantsupps1,by="Participant") %>%
  rename(bp_prop_wk3 = bp_prop3) %>%
  select(starts_with("Participant"),
         starts_with("bp_prop_wk3"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
bp_brmpsprop <- merge(fecal1_brmpsprop,fecal3_brmpsprop,
                           by="Participant") %>%
                     select(-ends_with("x"), -ends_with("y"))

#Appending Differences column to acetate-BRMPS df in form: week3-week1
bp_brmps_diffprop <- bp_brmpsprop %>%
  mutate(bp_prop_diff = bp_prop_wk3 - bp_prop_wk1) %>%
  arrange(by = bp_prop_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
bp_wk1_vecprop <- as.vector(bp_brmps_diffprop$bp_prop_wk1)
bp_wk3_vecprop <- as.vector(bp_brmps_diffprop$bp_prop_wk3)
paired_bp_dfprop <- data.frame(Week1 = bp_wk1_vecprop, 
                                    Week3 = bp_wk3_vecprop)
bppairplotprop <- ggpaired(paired_bp_dfprop, cond1 = "Week1", 
                                cond2 = "Week3")
bppairplotprop$layers <- bppairplotprop$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(bp_brmpsprop$bp_prop_wk1), 
       as.vector(bp_brmpsprop$bp_prop_wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(bp_brmps_diffprop) # p-value = 0.0971. 
bppairplotprop
bpprop1x_barplot <- barplot(bp_brmps_diffprop$bp_prop_diff, ylab = "Participants", xlab = "BP Prop. Change (Week 3 - Week 1)", 
        main = "LODAAT 1x Daily BP. Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(bp_brmps_diffprop)
```

##Total SCFA Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 1xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_total <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate")) %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_total,participantsupps1,by="Participant") %>%
  rename(total.wk1 = scfa_sum) %>%
  select(starts_with("Participant"),
        starts_with("total"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_total <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate")) %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_total,participantsupps1,by="Participant") %>%
  rename(total.wk3 = scfa_sum) %>%
  select(starts_with("Participant"),
         starts_with("total"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
total_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
total_brmps_diff <- total_brmps %>%
  mutate(total_diff = total.wk3 - total.wk1) %>%
  arrange(by = total_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
total_wk1_vec <- as.vector(total_brmps$total.wk1)
total_wk3_vec <- as.vector(total_brmps$total.wk3)
paired_total_df <- data.frame(Week1 = total_wk1_vec, Week3 = total_wk3_vec)
totalpairplot <- ggpaired(paired_total_df, cond1 = "Week1", cond2 = "Week3")
totalpairplot$layers <- totalpairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(total_brmps$total.wk1), 
       as.vector(total_brmps$total.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(total_brmps_diff)
totalpairplot
totbrmps1x_barplot <- barplot(total_brmps_diff$total_diff, ylab = "Participants", xlab = "Total Conc. Change (Week 3 - Week 1)", 
        main = "LODAAT 1x Daily Total Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(total_brmps_diff)
```

##pH Change in BRMPS 2x Cohort from Week 1 to Week 3
```{r}
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015") %>%
  rename(Participant = Name) %>%
  filter(Supplement.consumed == "LODAAT 1xdaily 20")

#Making Week 1 BRMPS 2x-pH Data Frame 
ph_brmps_wk1 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week1") %>%
  na.omit() %>%
  rename(ph.wk1 = pH)

#Making Week 3 BRMPS 2x-pH Data Frame 
ph_brmps_wk3 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week3") %>%
  na.omit() %>%
  rename(ph.wk3 = pH)

#Combining Week 1 and Week 3 BRMPS 2x Daily Cohort 
merged_ph_wks <- merge(ph_brmps_wk1, ph_brmps_wk3, by = "Participant") %>%
  select(-starts_with("Week"))
merged_ph_brmps<- merge(merged_ph_wks, participantsupps1, by = "Participant")
merged_ph_brmps <- merged_ph_brmps %>%
  select(starts_with("Participant"), starts_with("ph.wk")) %>%
  mutate(ph.diff = ph.wk3 - ph.wk1) %>%
  arrange(by_group = ph.diff)

#Summary Statistics and Visualizations for BRMPS 2x Daily pH Change
nrow(merged_ph_brmps) 
t.test(as.vector(merged_ph_brmps$ph.wk1), 
       as.vector(merged_ph_brmps$ph.wk3),
       alternative = "greater",
       paired = TRUE,
       conf.level = 0.95)
summary(merged_ph_brmps)
ph_brmps1x_barplot <- barplot(merged_ph_brmps$ph.diff, ylab = "Participants", 
        xlab = "pH (Week 3 - Week 1)", 
        main = "LODAAT 1x Daily pH Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(merged_ph_brmps)
```
#-------------------------------------------------------------------------------
#LODAAT 2x Daily 9
##Acetate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
fullfecalsample <- read.csv("./fecalsample_a.csv")

#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 9") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_acetate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_acetate,participantsupps1,by="Participant") %>%
  rename(acetate.wk1 = Acetate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("acetate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_acetate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate")) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_acetate,participantsupps1,by="Participant") %>%
  rename(acetate.wk3 = Acetate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("acetate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
acetate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
acetate_brmps_diff <- acetate_brmps %>%
  mutate(acetate_diff = acetate.wk3 - acetate.wk1) %>%
  arrange(by = acetate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
acetate_wk1_vec <- as.vector(acetate_brmps$acetate.wk1)
acetate_wk3_vec <- as.vector(acetate_brmps$acetate.wk3)
paired_acetate_df <- data.frame(Week1 = acetate_wk1_vec, Week3 = acetate_wk3_vec)
acetatepairplot <- ggpaired(paired_acetate_df, cond1 = "Week1", cond2 = "Week3")
acetatepairplot$layers <- acetatepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(acetate_brmps$acetate.wk1), 
       as.vector(acetate_brmps$acetate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(acetate_brmps_diff)
acetatepairplot
acebrmps1x_barplot <- barplot(acetate_brmps_diff$acetate_diff, ylab = "Participants", xlab = "Acet. Conc. Change (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 9 Acetate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(acetate_brmps_diff)
```

##Acetate Proportion in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 9") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_acetateprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week1") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(acetate_prop1 = Acetate_mmol_kg / scfa_sum) 

fecal1_brmpsprop <- merge(fecal1_acetateprop,participantsupps1,by="Participant") %>%
  rename(acetate_prop_wk1 = acetate_prop1) %>%
  select(starts_with("Participant"),
         starts_with("acetate_prop_wk1"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_acetateprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week3") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(acetate_prop3 = Acetate_mmol_kg / scfa_sum) 

fecal3_brmpsprop <- merge(fecal3_acetateprop,participantsupps1,by="Participant") %>%
  rename(acetate_prop_wk3 = acetate_prop3) %>%
  select(starts_with("Participant"),
         starts_with("acetate_prop_wk3"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
acetate_brmpsprop <- merge(fecal1_brmpsprop,fecal3_brmpsprop,
                           by="Participant") %>%
                     select(-ends_with("x"), -ends_with("y"))

#Appending Differences column to acetate-BRMPS df in form: week3-week1
acetate_brmps_diffprop <- acetate_brmpsprop %>%
  mutate(acetate_prop_diff = acetate_prop_wk3 - acetate_prop_wk1) %>%
  arrange(by = acetate_prop_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
acetate_wk1_vecprop <- as.vector(acetate_brmps_diffprop$acetate_prop_wk1)
acetate_wk3_vecprop <- as.vector(acetate_brmps_diffprop$acetate_prop_wk3)
paired_acetate_dfprop <- data.frame(Week1 = acetate_wk1_vecprop, 
                                    Week3 = acetate_wk3_vecprop)
acetatepairplotprop <- ggpaired(paired_acetate_dfprop, cond1 = "Week1", 
                                cond2 = "Week3")
acetatepairplotprop$layers <- acetatepairplotprop$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(acetate_brmpsprop$acetate_prop_wk1), 
       as.vector(acetate_brmpsprop$acetate_prop_wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(acetate_brmps_diffprop) # p-value = 0.0971. 
acetatepairplotprop
aceprop1x_barplot <- barplot(acetate_brmps_diffprop$acetate_prop_diff, ylab = "Participants", xlab = "Acetate Prop. Change (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 9 Ace. Prop. Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(acetate_brmps_diffprop)
```

##Butyrate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
fullfecalsample <- read.csv("./fecalsample_a.csv")

#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015") 

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 9") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_butyrate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Butyrate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_butyrate,participantsupps1,by="Participant") %>%
  rename(butyrate.wk1 = Butyrate_mmol_kg) %>%
  select(starts_with("Participant"),
        starts_with("butyrate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_butyrate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Butyrate")) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_butyrate,participantsupps1,by="Participant") %>%
  rename(butyrate.wk3 = Butyrate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("butyrate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
butyrate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round) 

#Appending Differences column to acetate-BRMPS df in form: week3-week1
butyrate_brmps_diff <- butyrate_brmps %>%
  mutate(butyrate_diff = butyrate.wk3 - butyrate.wk1) %>%
  arrange(by = butyrate_diff)

butyrate_brmps_diff <- butyrate_brmps_diff[-c(86, 87, 88), ]

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
butyrate_wk1_vec <- as.vector(butyrate_brmps$butyrate.wk1)
butyrate_wk3_vec <- as.vector(butyrate_brmps$butyrate.wk3)
paired_butyrate_df <- data.frame(Week1 = butyrate_wk1_vec, Week3 = butyrate_wk3_vec)
butyratepairplot <- ggpaired(paired_butyrate_df, cond1 = "Week1", cond2 = "Week3")
butyratepairplot$layers <- butyratepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(butyrate_brmps_diff$butyrate.wk1), 
       as.vector(butyrate_brmps_diff$butyrate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(butyrate_brmps_diff)
butyratepairplot
butbrmps1x_barplot <- barplot(butyrate_brmps_diff$butyrate_diff, ylab = "Participants", xlab = "Butyr. Conc. Change (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 9 Butyrate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(butyrate_brmps_diff)
```

##Propionate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 9") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_propionate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Propionate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_propionate,participantsupps1,by="Participant") %>%
  rename(propionate.wk1 = Propionate_mmol_kg) %>%
  select(starts_with("Participant"),
        starts_with("propionate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_propionate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Propionate")) %>%
  filter(Week == "week3") %>%
  na.omit()

fecal3_brmps <- merge(fecal3_propionate,participantsupps1,by="Participant") %>%
  rename(propionate.wk3 = Propionate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("propionate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
propionate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
propionate_brmps_diff <- propionate_brmps %>%
  mutate(propionate_diff = propionate.wk3 - propionate.wk1) %>%
  arrange(by = propionate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
propionate_wk1_vec <- as.vector(propionate_brmps$propionate.wk1)
propionate_wk3_vec <- as.vector(propionate_brmps$propionate.wk3)
paired_propionate_df <- data.frame(Week1 = propionate_wk1_vec, Week3 = propionate_wk3_vec)
propionatepairplot <- ggpaired(paired_propionate_df, cond1 = "Week1", cond2 = "Week3")
propionatepairplot$layers <- propionatepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(propionate_brmps$propionate.wk1), 
       as.vector(propionate_brmps$propionate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(propionate_brmps_diff)
propionatepairplot
probrmps1x_barplot <- barplot(propionate_brmps_diff$propionate_diff, ylab = "Participants", xlab = "Propio. Conc. Change (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 9 Propionate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(propionate_brmps_diff)
```
##Prop+But Proportion (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 9") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_bpprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week1") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(bp_prop1 = (Butyrate_mmol_kg+Propionate_mmol_kg) / scfa_sum) 
#View(fullfecalsample)

fecal1_brmpsprop <- merge(fecal1_bpprop,participantsupps1,by="Participant") %>%
  rename(bp_prop_wk1 = bp_prop1) %>%
  select(starts_with("Participant"),
         starts_with("bp_prop_wk1"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_bpprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week3") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(bp_prop3 = (Butyrate_mmol_kg+Propionate_mmol_kg) / scfa_sum) 

fecal3_brmpsprop <- merge(fecal3_bpprop,participantsupps1,by="Participant") %>%
  rename(bp_prop_wk3 = bp_prop3) %>%
  select(starts_with("Participant"),
         starts_with("bp_prop_wk3"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
bp_brmpsprop <- merge(fecal1_brmpsprop,fecal3_brmpsprop,
                           by="Participant") %>%
                     select(-ends_with("x"), -ends_with("y"))

#Appending Differences column to acetate-BRMPS df in form: week3-week1
bp_brmps_diffprop <- bp_brmpsprop %>%
  mutate(bp_prop_diff = bp_prop_wk3 - bp_prop_wk1) %>%
  arrange(by = bp_prop_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
bp_wk1_vecprop <- as.vector(bp_brmps_diffprop$bp_prop_wk1)
bp_wk3_vecprop <- as.vector(bp_brmps_diffprop$bp_prop_wk3)
paired_bp_dfprop <- data.frame(Week1 = bp_wk1_vecprop, 
                                    Week3 = bp_wk3_vecprop)
bppairplotprop <- ggpaired(paired_bp_dfprop, cond1 = "Week1", 
                                cond2 = "Week3")
bppairplotprop$layers <- bppairplotprop$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(bp_brmpsprop$bp_prop_wk1), 
       as.vector(bp_brmpsprop$bp_prop_wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(bp_brmps_diffprop) # p-value = 0.0971. 
bppairplotprop
bpprop1x_barplot <- barplot(bp_brmps_diffprop$bp_prop_diff, ylab = "Participants", xlab = "BP Prop. Change (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 9 BP. Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(bp_brmps_diffprop)
```

##Total SCFA Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 9") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_total <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate")) %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_total,participantsupps1,by="Participant") %>%
  rename(total.wk1 = scfa_sum) %>%
  select(starts_with("Participant"),
        starts_with("total"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_total <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate")) %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_total,participantsupps1,by="Participant") %>%
  rename(total.wk3 = scfa_sum) %>%
  select(starts_with("Participant"),
         starts_with("total"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
total_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
total_brmps_diff <- total_brmps %>%
  mutate(total_diff = total.wk3 - total.wk1) %>%
  arrange(by = total_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
total_wk1_vec <- as.vector(total_brmps$total.wk1)
total_wk3_vec <- as.vector(total_brmps$total.wk3)
paired_total_df <- data.frame(Week1 = total_wk1_vec, Week3 = total_wk3_vec)
totalpairplot <- ggpaired(paired_total_df, cond1 = "Week1", cond2 = "Week3")
totalpairplot$layers <- totalpairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(total_brmps$total.wk1), 
       as.vector(total_brmps$total.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(total_brmps_diff)
totalpairplot
totbrmps1x_barplot <- barplot(total_brmps_diff$total_diff, ylab = "Participants", xlab = "Total Conc. Change (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 9 Total Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(total_brmps_diff)
```

##pH Change in BRMPS 2x Cohort from Week 1 to Week 3
```{r}
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015") %>%
  rename(Participant = Name) %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 9")

#Making Week 1 BRMPS 2x-pH Data Frame 
ph_brmps_wk1 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week1") %>%
  na.omit() %>%
  rename(ph.wk1 = pH)

#Making Week 3 BRMPS 2x-pH Data Frame 
ph_brmps_wk3 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week3") %>%
  na.omit() %>%
  rename(ph.wk3 = pH)

#Combining Week 1 and Week 3 BRMPS 2x Daily Cohort 
merged_ph_wks <- merge(ph_brmps_wk1, ph_brmps_wk3, by = "Participant") %>%
  select(-starts_with("Week"))
merged_ph_brmps<- merge(merged_ph_wks, participantsupps1, by = "Participant")
merged_ph_brmps <- merged_ph_brmps %>%
  select(starts_with("Participant"), starts_with("ph.wk")) %>%
  mutate(ph.diff = ph.wk3 - ph.wk1) %>%
  arrange(by_group = ph.diff)

#Summary Statistics and Visualizations for BRMPS 2x Daily pH Change
nrow(merged_ph_brmps) 
t.test(as.vector(merged_ph_brmps$ph.wk1), 
       as.vector(merged_ph_brmps$ph.wk3),
       alternative = "greater",
       paired = TRUE,
       conf.level = 0.95)
summary(merged_ph_brmps)
ph_brmps1x_barplot <- barplot(merged_ph_brmps$ph.diff, ylab = "Participants", 
        xlab = "pH (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 9 pH Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(merged_ph_brmps)
```
#-------------------------------------------------------------------------------
#LODAAT 2x Daily 18 and 20
##Acetate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
fullfecalsample <- read.csv("./fecalsample_a.csv")

#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 18" | Supplement.consumed == "LODAAT 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_acetate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_acetate,participantsupps1,by="Participant") %>%
  rename(acetate.wk1 = Acetate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("acetate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_acetate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate")) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_acetate,participantsupps1,by="Participant") %>%
  rename(acetate.wk3 = Acetate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("acetate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
acetate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
acetate_brmps_diff <- acetate_brmps %>%
  mutate(acetate_diff = acetate.wk3 - acetate.wk1) %>%
  arrange(by = acetate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
acetate_wk1_vec <- as.vector(acetate_brmps$acetate.wk1)
acetate_wk3_vec <- as.vector(acetate_brmps$acetate.wk3)
paired_acetate_df <- data.frame(Week1 = acetate_wk1_vec, Week3 = acetate_wk3_vec)
acetatepairplot <- ggpaired(paired_acetate_df, cond1 = "Week1", cond2 = "Week3")
acetatepairplot$layers <- acetatepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(acetate_brmps$acetate.wk1), 
       as.vector(acetate_brmps$acetate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(acetate_brmps_diff)
acetatepairplot
acebrmps1x_barplot <- barplot(acetate_brmps_diff$acetate_diff, ylab = "Participants", xlab = "Acet. Conc. Change (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 18/20 Acetate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(acetate_brmps_diff)
```

##Acetate Proportion in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 18" | Supplement.consumed == "LODAAT 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_acetateprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week1") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(acetate_prop1 = Acetate_mmol_kg / scfa_sum) 

fecal1_brmpsprop <- merge(fecal1_acetateprop,participantsupps1,by="Participant") %>%
  rename(acetate_prop_wk1 = acetate_prop1) %>%
  select(starts_with("Participant"),
         starts_with("acetate_prop_wk1"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_acetateprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week3") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(acetate_prop3 = Acetate_mmol_kg / scfa_sum) 

fecal3_brmpsprop <- merge(fecal3_acetateprop,participantsupps1,by="Participant") %>%
  rename(acetate_prop_wk3 = acetate_prop3) %>%
  select(starts_with("Participant"),
         starts_with("acetate_prop_wk3"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
acetate_brmpsprop <- merge(fecal1_brmpsprop,fecal3_brmpsprop,
                           by="Participant") %>%
                     select(-ends_with("x"), -ends_with("y"))

#Appending Differences column to acetate-BRMPS df in form: week3-week1
acetate_brmps_diffprop <- acetate_brmpsprop %>%
  mutate(acetate_prop_diff = acetate_prop_wk3 - acetate_prop_wk1) %>%
  arrange(by = acetate_prop_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
acetate_wk1_vecprop <- as.vector(acetate_brmps_diffprop$acetate_prop_wk1)
acetate_wk3_vecprop <- as.vector(acetate_brmps_diffprop$acetate_prop_wk3)
paired_acetate_dfprop <- data.frame(Week1 = acetate_wk1_vecprop, 
                                    Week3 = acetate_wk3_vecprop)
acetatepairplotprop <- ggpaired(paired_acetate_dfprop, cond1 = "Week1", 
                                cond2 = "Week3")
acetatepairplotprop$layers <- acetatepairplotprop$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(acetate_brmpsprop$acetate_prop_wk1), 
       as.vector(acetate_brmpsprop$acetate_prop_wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(acetate_brmps_diffprop) # p-value = 0.0971. 
acetatepairplotprop
aceprop1x_barplot <- barplot(acetate_brmps_diffprop$acetate_prop_diff, ylab = "Participants", xlab = "Acetate Prop. Change (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 18/20 Ace. Prop. Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(acetate_brmps_diffprop)
```

##Butyrate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
fullfecalsample <- read.csv("./fecalsample_a.csv")

#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015") 

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 18" | Supplement.consumed == "LODAAT 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_butyrate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Butyrate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_butyrate,participantsupps1,by="Participant") %>%
  rename(butyrate.wk1 = Butyrate_mmol_kg) %>%
  select(starts_with("Participant"),
        starts_with("butyrate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_butyrate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Butyrate")) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_butyrate,participantsupps1,by="Participant") %>%
  rename(butyrate.wk3 = Butyrate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("butyrate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
butyrate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round) 

#Appending Differences column to acetate-BRMPS df in form: week3-week1
butyrate_brmps_diff <- butyrate_brmps %>%
  mutate(butyrate_diff = butyrate.wk3 - butyrate.wk1) %>%
  arrange(by = butyrate_diff)

butyrate_brmps_diff <- butyrate_brmps_diff[-c(86, 87, 88), ]

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
butyrate_wk1_vec <- as.vector(butyrate_brmps$butyrate.wk1)
butyrate_wk3_vec <- as.vector(butyrate_brmps$butyrate.wk3)
paired_butyrate_df <- data.frame(Week1 = butyrate_wk1_vec, Week3 = butyrate_wk3_vec)
butyratepairplot <- ggpaired(paired_butyrate_df, cond1 = "Week1", cond2 = "Week3")
butyratepairplot$layers <- butyratepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(butyrate_brmps_diff$butyrate.wk1), 
       as.vector(butyrate_brmps_diff$butyrate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(butyrate_brmps_diff)
butyratepairplot
butbrmps1x_barplot <- barplot(butyrate_brmps_diff$butyrate_diff, ylab = "Participants", xlab = "Butyr. Conc. Change (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 18/20 Butyrate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(butyrate_brmps_diff)
```

##Propionate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 18" | Supplement.consumed == "LODAAT 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_propionate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Propionate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_propionate,participantsupps1,by="Participant") %>%
  rename(propionate.wk1 = Propionate_mmol_kg) %>%
  select(starts_with("Participant"),
        starts_with("propionate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_propionate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Propionate")) %>%
  filter(Week == "week3") %>%
  na.omit()

fecal3_brmps <- merge(fecal3_propionate,participantsupps1,by="Participant") %>%
  rename(propionate.wk3 = Propionate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("propionate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
propionate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
propionate_brmps_diff <- propionate_brmps %>%
  mutate(propionate_diff = propionate.wk3 - propionate.wk1) %>%
  arrange(by = propionate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
propionate_wk1_vec <- as.vector(propionate_brmps$propionate.wk1)
propionate_wk3_vec <- as.vector(propionate_brmps$propionate.wk3)
paired_propionate_df <- data.frame(Week1 = propionate_wk1_vec, Week3 = propionate_wk3_vec)
propionatepairplot <- ggpaired(paired_propionate_df, cond1 = "Week1", cond2 = "Week3")
propionatepairplot$layers <- propionatepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(propionate_brmps$propionate.wk1), 
       as.vector(propionate_brmps$propionate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(propionate_brmps_diff)
propionatepairplot
probrmps1x_barplot <- barplot(propionate_brmps_diff$propionate_diff, ylab = "Participants", xlab = "Propio. Conc. Change (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 18/20 Propionate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(propionate_brmps_diff)
```
##Prop+But Proportion (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 18" | Supplement.consumed == "LODAAT 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_bpprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week1") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(bp_prop1 = (Butyrate_mmol_kg+Propionate_mmol_kg) / scfa_sum) 
#View(fullfecalsample)

fecal1_brmpsprop <- merge(fecal1_bpprop,participantsupps1,by="Participant") %>%
  rename(bp_prop_wk1 = bp_prop1) %>%
  select(starts_with("Participant"),
         starts_with("bp_prop_wk1"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_bpprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week3") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(bp_prop3 = (Butyrate_mmol_kg+Propionate_mmol_kg) / scfa_sum) 

fecal3_brmpsprop <- merge(fecal3_bpprop,participantsupps1,by="Participant") %>%
  rename(bp_prop_wk3 = bp_prop3) %>%
  select(starts_with("Participant"),
         starts_with("bp_prop_wk3"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
bp_brmpsprop <- merge(fecal1_brmpsprop,fecal3_brmpsprop,
                           by="Participant") %>%
                     select(-ends_with("x"), -ends_with("y"))

#Appending Differences column to acetate-BRMPS df in form: week3-week1
bp_brmps_diffprop <- bp_brmpsprop %>%
  mutate(bp_prop_diff = bp_prop_wk3 - bp_prop_wk1) %>%
  arrange(by = bp_prop_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
bp_wk1_vecprop <- as.vector(bp_brmps_diffprop$bp_prop_wk1)
bp_wk3_vecprop <- as.vector(bp_brmps_diffprop$bp_prop_wk3)
paired_bp_dfprop <- data.frame(Week1 = bp_wk1_vecprop, 
                                    Week3 = bp_wk3_vecprop)
bppairplotprop <- ggpaired(paired_bp_dfprop, cond1 = "Week1", 
                                cond2 = "Week3")
bppairplotprop$layers <- bppairplotprop$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(bp_brmpsprop$bp_prop_wk1), 
       as.vector(bp_brmpsprop$bp_prop_wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(bp_brmps_diffprop) # p-value = 0.0971. 
bppairplotprop
bpprop1x_barplot <- barplot(bp_brmps_diffprop$bp_prop_diff, ylab = "Participants", xlab = "BP Prop. Change (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 18/20 BP. Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(bp_brmps_diffprop)
```

##Total SCFA Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 18" | Supplement.consumed == "LODAAT 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_total <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate")) %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_total,participantsupps1,by="Participant") %>%
  rename(total.wk1 = scfa_sum) %>%
  select(starts_with("Participant"),
        starts_with("total"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_total <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate")) %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_total,participantsupps1,by="Participant") %>%
  rename(total.wk3 = scfa_sum) %>%
  select(starts_with("Participant"),
         starts_with("total"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
total_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
total_brmps_diff <- total_brmps %>%
  mutate(total_diff = total.wk3 - total.wk1) %>%
  arrange(by = total_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
total_wk1_vec <- as.vector(total_brmps$total.wk1)
total_wk3_vec <- as.vector(total_brmps$total.wk3)
paired_total_df <- data.frame(Week1 = total_wk1_vec, Week3 = total_wk3_vec)
totalpairplot <- ggpaired(paired_total_df, cond1 = "Week1", cond2 = "Week3")
totalpairplot$layers <- totalpairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(total_brmps$total.wk1), 
       as.vector(total_brmps$total.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(total_brmps_diff)
totalpairplot
totbrmps1x_barplot <- barplot(total_brmps_diff$total_diff, ylab = "Participants", xlab = "Total Conc. Change (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 18/20 Total Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(total_brmps_diff)
```

##pH Change in BRMPS 2x Cohort from Week 1 to Week 3
```{r}
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015") %>%
  rename(Participant = Name) %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 18" | Supplement.consumed == "LODAAT 2xdaily 20")

fullfecalsample_unaveraged<- fullfecalsample_unaveraged %>%
   mutate(scfa_sum = Acetate_mmol_kg+Butyrate_mmol_kg+Propionate_mmol_kg) 

#Making Week 1 BRMPS 2x-pH Data Frame 
ph_brmps_wk1 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week1") %>%
  na.omit() %>%
  rename(ph.wk1 = pH)

#Making Week 3 BRMPS 2x-pH Data Frame 
ph_brmps_wk3 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week3") %>%
  na.omit() %>%
  rename(ph.wk3 = pH)

#Combining Week 1 and Week 3 BRMPS 2x Daily Cohort 
merged_ph_wks <- merge(ph_brmps_wk1, ph_brmps_wk3, by = "Participant") %>%
  select(-starts_with("Week"))
merged_ph_brmps<- merge(merged_ph_wks, participantsupps1, by = "Participant")
merged_ph_brmps <- merged_ph_brmps %>%
  select(starts_with("Participant"), starts_with("ph.wk")) %>%
  mutate(ph.diff = ph.wk3 - ph.wk1) %>%
  arrange(by_group = ph.diff)
#Summary Statistics and Visualizations for BRMPS 2x Daily pH Change
nrow(merged_ph_brmps) 
t.test(as.vector(merged_ph_brmps$ph.wk1), 
       as.vector(merged_ph_brmps$ph.wk3),
       alternative = "greater",
       paired = TRUE,
       conf.level = 0.95)
summary(merged_ph_brmps)
ph_brmps1x_barplot <- barplot(merged_ph_brmps$ph.diff, ylab = "Participants", 
        xlab = "pH (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 18/20 pH Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(merged_ph_brmps)
```

#-------------------------------------------------------------------------------
#Semester ANOVA Analyses
##ANOVA: Week 1 Total SCFA By Semester
```{r}
#Reading in Fecal Sample Data-Large Set
fullfecalsample <- read.csv("./fecalsample_a.csv") 
participants <- read.csv("./participant.csv")

#Making Table with Relevant Semester Info. Only
participants <- participants %>%
  filter(Data.consent == "True", Semester != "Winter2015") %>%
  filter(Quantity.compliant == "yes" | Quantity.compliant == "none") %>%
  rename(Participant = Name) %>%
  select(Participant, Semester) 

#Filtering for Week 1 and Merging Semester Info. by Participant
week1sample <- fullfecalsample %>%
  filter(Week == "week1") %>%
  select(Participant, Week, Acetate_mmol_kg, Butyrate_mmol_kg, Propionate_mmol_kg)

week1sample <- merge(participants, week1sample, by = "Participant") %>%
  mutate(scfa_sum = rowSums(.[4:6])) %>%
  filter(Semester != "Fall2017") %>%
  na.omit()

#Tukey Intervals to find Different Groups
summary(aov(scfa_sum ~ Semester, data = week1sample))
fit <- aov(scfa_sum ~ Semester, data = week1sample)
TukeyHSD(fit)

#Making Jitter Plot
e <- ggplot(week1sample, aes(x = Semester, y = scfa_sum))
e + geom_jitter(
  aes(color = Semester), 
  position = position_jitter(0.2),
  size = 1.2
  ) +
  stat_summary(
    aes(color = Semester),
    fun.data="mean_sdl",  fun.args = list(mult=1), 
    geom = "errorbar",  size = 0.4, color = "black"
  ) +
  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  coord_cartesian(ylim=c(0,175))

#Making Bargraph for scfa_sum by Semester 
week1sample <- week1sample %>%
  na.omit()

week1sample <- aggregate(x = week1sample$scfa_sum,                
          by = list(week1sample$Semester),          
          FUN = mean)

#Plot the bar chart 
week1samplebarplot<- barplot(as.vector(week1sample$x),
                             names.arg=as.vector(week1sample$Group.1), xlab="Semester",ylab="Mean Total SCFA (mmol/kg)",col="black",
main="Average Total SCFA by Semester",border="red", las=2,cex.names=0.5)

#Produce a measure of variance as well. 
```

##ANOVA: Week 1 Acetate By Semester
```{r}
#Reading in Fecal Sample Data-Large Set
fullfecalsample <- read.csv("./fecalsample_a.csv") 
participants <- read.csv("./participant.csv")

#Making Table with Relevant Semester Info. Only
participants <- participants %>%
  filter(Data.consent == "True", Semester != "Winter2015") %>%
  filter(Quantity.compliant == "yes" | Quantity.compliant == "none") %>%
  rename(Participant = Name) %>%
  select(Participant, Semester) 

#Filtering for Week 1 and Merging Semester Info. by Participant
week1sample <- fullfecalsample %>%
  filter(Week == "week1") %>%
  select(Participant, Week, Acetate_mmol_kg, Butyrate_mmol_kg, Propionate_mmol_kg)

##Tukey Intervals to find Different Groups
week1sample <- merge(participants, week1sample, by = "Participant") %>%
  mutate(scfa_sum = rowSums(.[4:6])) %>%
  filter(Semester != "Fall2017") %>%
  na.omit()
fit <- aov(Acetate_mmol_kg ~ Semester, data = week1sample)
summary(fit)
TukeyHSD(fit)

#Making Jitter Plot
e <- ggplot(week1sample, aes(x = Semester, y = Acetate_mmol_kg))
e + geom_jitter(
  aes(color = Semester), 
  position = position_jitter(0.2),
  size = 1.2
  ) +
  stat_summary(
    aes(color = Semester),
    fun.data="mean_sdl",  fun.args = list(mult=1), 
    geom = "errorbar",  size = 0.4, color = "black"
  ) +
  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  coord_cartesian(ylim=c(0,110))


#Making Bargraph for Acetate by Semester 
week1sample <- week1sample %>%
  na.omit()

week1sample <- aggregate(x = week1sample$Acetate_mmol_kg,                
          by = list(week1sample$Semester),          
          FUN = mean)

#Plot the bar chart 
week1samplebarplot<- barplot(as.vector(week1sample$x),
                             names.arg=as.vector(week1sample$Group.1), xlab="Semester",ylab="Mean Acetate (mmol/kg)",col="black",
main="Average Acetate by Semester",border="red", las=2,cex.names=0.5)
```

##ANOVA: Week 1 Butyrate By Semester
```{r}
#Reading in Fecal Sample Data-Large Set
fullfecalsample <- read.csv("./fecalsample_a.csv") 
participants <- read.csv("./participant.csv")

#Making Table with Relevant Semester Info. Only
participants <- participants %>%
  filter(Data.consent == "True", Semester != "Winter2015") %>%
  filter(Quantity.compliant == "yes" | Quantity.compliant == "none") %>%
  rename(Participant = Name) %>%
  select(Participant, Semester) 

#Filtering for Week 1 and Merging Semester Info. by Participant
week1sample <- fullfecalsample %>%
  filter(Week == "week1") %>%
  select(Participant, Week, Acetate_mmol_kg, Butyrate_mmol_kg, Propionate_mmol_kg)

#Tukey Intervals to find Different Groups
week1sample <- merge(participants, week1sample, by = "Participant") %>%
  mutate(scfa_sum = rowSums(.[4:6])) %>%
  filter(Semester != "Fall2017") %>%
  na.omit()
fit <- aov(Butyrate_mmol_kg ~ Semester, data = week1sample)
summary(fit)
TukeyHSD(fit)

#Making Jitter Plot
e <- ggplot(week1sample, aes(x = Semester, y = Butyrate_mmol_kg))
e + geom_jitter(
  aes(color = Semester), 
  position = position_jitter(0.2),
  size = 1.2
  ) +
  stat_summary(
    aes(color = Semester),
    fun.data="mean_sdl",  fun.args = list(mult=1), 
    geom = "errorbar",  size = 0.4, color = "black"
  ) +
  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  coord_cartesian(ylim=c(0,40))


#Making Bargraph for Butyrate by Semester 
week1sample <- week1sample %>%
  na.omit()

week1sample <- aggregate(x = week1sample$Butyrate_mmol_kg,                
          by = list(week1sample$Semester),          
          FUN = mean)

#Plot the bar chart 
week1samplebarplot<- barplot(as.vector(week1sample$x),
                             names.arg=as.vector(week1sample$Group.1), xlab="Semester",ylab="Mean Butyrate (mmol/kg)",col="black",
main="Average Butyrate by Semester",border="red", las=2,cex.names=0.5)
```

##ANOVA: Week 1 Propionate By Semester
```{r}
#Reading in Fecal Sample Data-Large Set
fullfecalsample <- read.csv("./fecalsample_a.csv") 
participants <- read.csv("./participant.csv")

#Making Table with Relevant Semester Info. Only
participants <- participants %>%
  filter(Data.consent == "True", Semester != "Winter2015") %>%
  filter(Quantity.compliant == "yes" | Quantity.compliant == "none") %>%
  rename(Participant = Name) %>%
  select(Participant, Semester) 

#Filtering for Week 1 and Merging Semester Info. by Participant
week1sample <- fullfecalsample %>%
  filter(Week == "week1") %>%
  select(Participant, Week, Acetate_mmol_kg, Butyrate_mmol_kg, Propionate_mmol_kg)

#Tukey Intervals to find Different Groups
week1sample <- merge(participants, week1sample, by = "Participant") %>%
  mutate(scfa_sum = rowSums(.[4:6])) %>%
  filter(Semester != "Fall2017") %>%
  na.omit()
fit <- aov(Propionate_mmol_kg ~ Semester, data = week1sample)
summary(fit)
TukeyHSD(fit)

#Making Jitter Plot
e <- ggplot(week1sample, aes(x = Semester, y = Propionate_mmol_kg))
e + geom_jitter(
  aes(color = Semester), 
  position = position_jitter(0.2),
  size = 1.2
  ) +
  stat_summary(
    aes(color = Semester),
    fun.data="mean_sdl",  fun.args = list(mult=1), 
    geom = "errorbar",  size = 0.4, color = "black"
  ) +
  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  coord_cartesian(ylim=c(0,40))


#Making Bargraph for Propionate by Semester 
week1sample <- week1sample %>%
  na.omit()

week1sample <- aggregate(x = week1sample$Propionate_mmol_kg,                
          by = list(week1sample$Semester),          
          FUN = mean)

#Plot the bar chart 
week1samplebarplot<- barplot(as.vector(week1sample$x),
                             names.arg=as.vector(week1sample$Group.1), xlab="Semester",ylab="Mean Propionate (mmol/kg)",col="black",
main="Average Propionate by Semester",border="red", las=2,cex.names=0.5)
```

#-------------------------------------------------------------------------------
#BRMPS2x20 + LODAAT18/20 
##Acetate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
fullfecalsample <- read.csv("./fecalsample_a.csv")

#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 18" | 
           Supplement.consumed == "LODAAT 2xdaily 20" | 
           Supplement.consumed == "BRMPS 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_acetate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_acetate,participantsupps1,by="Participant") %>%
  rename(acetate.wk1 = Acetate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("acetate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_acetate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate")) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_acetate,participantsupps1,by="Participant") %>%
  rename(acetate.wk3 = Acetate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("acetate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
acetate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
acetate_brmps_diff <- acetate_brmps %>%
  mutate(acetate_diff = acetate.wk3 - acetate.wk1) %>%
  arrange(by = acetate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
acetate_wk1_vec <- as.vector(acetate_brmps$acetate.wk1)
acetate_wk3_vec <- as.vector(acetate_brmps$acetate.wk3)
paired_acetate_df <- data.frame(Week1 = acetate_wk1_vec, Week3 = acetate_wk3_vec)
acetatepairplot <- ggpaired(paired_acetate_df, cond1 = "Week1", cond2 = "Week3")
acetatepairplot$layers <- acetatepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(acetate_brmps$acetate.wk1), 
       as.vector(acetate_brmps$acetate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(acetate_brmps_diff)
acetatepairplot
acebrmps1x_barplot <- barplot(acetate_brmps_diff$acetate_diff, ylab = "Participants", xlab = "Acet. Conc. Change (Week 3 - Week 1)", 
        main = "BRMPS 2x 20 + LODAAT 2x Daily 18/20 Acetate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(acetate_brmps_diff)
```

##Acetate Proportion in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 18" | Supplement.consumed == "LODAAT 2xdaily 20" | Supplement.consumed == "BRMPS 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_acetateprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week1") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(acetate_prop1 = Acetate_mmol_kg / scfa_sum) 

fecal1_brmpsprop <- merge(fecal1_acetateprop,participantsupps1,by="Participant") %>%
  rename(acetate_prop_wk1 = acetate_prop1) %>%
  select(starts_with("Participant"),
         starts_with("acetate_prop_wk1"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_acetateprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week3") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(acetate_prop3 = Acetate_mmol_kg / scfa_sum) 

fecal3_brmpsprop <- merge(fecal3_acetateprop,participantsupps1,by="Participant") %>%
  rename(acetate_prop_wk3 = acetate_prop3) %>%
  select(starts_with("Participant"),
         starts_with("acetate_prop_wk3"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
acetate_brmpsprop <- merge(fecal1_brmpsprop,fecal3_brmpsprop,
                           by="Participant") %>%
                     select(-ends_with("x"), -ends_with("y"))

#Appending Differences column to acetate-BRMPS df in form: week3-week1
acetate_brmps_diffprop <- acetate_brmpsprop %>%
  mutate(acetate_prop_diff = acetate_prop_wk3 - acetate_prop_wk1) %>%
  arrange(by = acetate_prop_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
acetate_wk1_vecprop <- as.vector(acetate_brmps_diffprop$acetate_prop_wk1)
acetate_wk3_vecprop <- as.vector(acetate_brmps_diffprop$acetate_prop_wk3)
paired_acetate_dfprop <- data.frame(Week1 = acetate_wk1_vecprop, 
                                    Week3 = acetate_wk3_vecprop)
acetatepairplotprop <- ggpaired(paired_acetate_dfprop, cond1 = "Week1", 
                                cond2 = "Week3")
acetatepairplotprop$layers <- acetatepairplotprop$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(acetate_brmpsprop$acetate_prop_wk1), 
       as.vector(acetate_brmpsprop$acetate_prop_wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(acetate_brmps_diffprop) # p-value = 0.0971. 
acetatepairplotprop
aceprop1x_barplot <- barplot(acetate_brmps_diffprop$acetate_prop_diff, ylab = "Participants", xlab = "Acetate Prop. Change (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 18/20 Ace. Prop. Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(acetate_brmps_diffprop)
```

##Butyrate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
fullfecalsample <- read.csv("./fecalsample_a.csv")

#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015") 

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 18" | Supplement.consumed == "LODAAT 2xdaily 20" | Supplement.consumed == "BRMPS 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_butyrate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Butyrate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_butyrate,participantsupps1,by="Participant") %>%
  rename(butyrate.wk1 = Butyrate_mmol_kg) %>%
  select(starts_with("Participant"),
        starts_with("butyrate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_butyrate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Butyrate")) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_butyrate,participantsupps1,by="Participant") %>%
  rename(butyrate.wk3 = Butyrate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("butyrate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
butyrate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round) 

#Appending Differences column to acetate-BRMPS df in form: week3-week1
butyrate_brmps_diff <- butyrate_brmps %>%
  mutate(butyrate_diff = butyrate.wk3 - butyrate.wk1) %>%
  arrange(by = butyrate_diff)

butyrate_brmps_diff <- butyrate_brmps_diff[-c(86, 87, 88), ]

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
butyrate_wk1_vec <- as.vector(butyrate_brmps$butyrate.wk1)
butyrate_wk3_vec <- as.vector(butyrate_brmps$butyrate.wk3)
paired_butyrate_df <- data.frame(Week1 = butyrate_wk1_vec, Week3 = butyrate_wk3_vec)
butyratepairplot <- ggpaired(paired_butyrate_df, cond1 = "Week1", cond2 = "Week3")
butyratepairplot$layers <- butyratepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(butyrate_brmps_diff$butyrate.wk1), 
       as.vector(butyrate_brmps_diff$butyrate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(butyrate_brmps_diff)
butyratepairplot
butbrmps1x_barplot <- barplot(butyrate_brmps_diff$butyrate_diff, ylab = "Participants", xlab = "Butyr. Conc. Change (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 18/20 Butyrate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(butyrate_brmps_diff)
```

##Propionate Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 18" | Supplement.consumed == "LODAAT 2xdaily 20" | Supplement.consumed == "BRMPS 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_propionate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Propionate")) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_propionate,participantsupps1,by="Participant") %>%
  rename(propionate.wk1 = Propionate_mmol_kg) %>%
  select(starts_with("Participant"),
        starts_with("propionate"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_propionate <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Propionate")) %>%
  filter(Week == "week3") %>%
  na.omit()

fecal3_brmps <- merge(fecal3_propionate,participantsupps1,by="Participant") %>%
  rename(propionate.wk3 = Propionate_mmol_kg) %>%
  select(starts_with("Participant"),
         starts_with("propionate"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
propionate_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
propionate_brmps_diff <- propionate_brmps %>%
  mutate(propionate_diff = propionate.wk3 - propionate.wk1) %>%
  arrange(by = propionate_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
propionate_wk1_vec <- as.vector(propionate_brmps$propionate.wk1)
propionate_wk3_vec <- as.vector(propionate_brmps$propionate.wk3)
paired_propionate_df <- data.frame(Week1 = propionate_wk1_vec, Week3 = propionate_wk3_vec)
propionatepairplot <- ggpaired(paired_propionate_df, cond1 = "Week1", cond2 = "Week3")
propionatepairplot$layers <- propionatepairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(propionate_brmps$propionate.wk1), 
       as.vector(propionate_brmps$propionate.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(propionate_brmps_diff)
propionatepairplot
probrmps1x_barplot <- barplot(propionate_brmps_diff$propionate_diff, ylab = "Participants", xlab = "Propio. Conc. Change (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 18/20 Propionate Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(propionate_brmps_diff)
```
##Prop+But Proportion (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 18" | Supplement.consumed == "LODAAT 2xdaily 20" | Supplement.consumed == "BRMPS 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_bpprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week1") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(bp_prop1 = (Butyrate_mmol_kg+Propionate_mmol_kg) / scfa_sum) 
#View(fullfecalsample)

fecal1_brmpsprop <- merge(fecal1_bpprop,participantsupps1,by="Participant") %>%
  rename(bp_prop_wk1 = bp_prop1) %>%
  select(starts_with("Participant"),
         starts_with("bp_prop_wk1"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_bpprop <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate"))%>%
  filter(Week == "week3") %>%
  na.omit() %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  mutate(bp_prop3 = (Butyrate_mmol_kg+Propionate_mmol_kg) / scfa_sum) 

fecal3_brmpsprop <- merge(fecal3_bpprop,participantsupps1,by="Participant") %>%
  rename(bp_prop_wk3 = bp_prop3) %>%
  select(starts_with("Participant"),
         starts_with("bp_prop_wk3"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
bp_brmpsprop <- merge(fecal1_brmpsprop,fecal3_brmpsprop,
                           by="Participant") %>%
                     select(-ends_with("x"), -ends_with("y"))

#Appending Differences column to acetate-BRMPS df in form: week3-week1
bp_brmps_diffprop <- bp_brmpsprop %>%
  mutate(bp_prop_diff = bp_prop_wk3 - bp_prop_wk1) %>%
  arrange(by = bp_prop_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
bp_wk1_vecprop <- as.vector(bp_brmps_diffprop$bp_prop_wk1)
bp_wk3_vecprop <- as.vector(bp_brmps_diffprop$bp_prop_wk3)
paired_bp_dfprop <- data.frame(Week1 = bp_wk1_vecprop, 
                                    Week3 = bp_wk3_vecprop)
bppairplotprop <- ggpaired(paired_bp_dfprop, cond1 = "Week1", 
                                cond2 = "Week3")
bppairplotprop$layers <- bppairplotprop$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(bp_brmpsprop$bp_prop_wk1), 
       as.vector(bp_brmpsprop$bp_prop_wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(bp_brmps_diffprop) # p-value = 0.0971. 
bppairplotprop
bpprop1x_barplot <- barplot(bp_brmps_diffprop$bp_prop_diff, ylab = "Participants", xlab = "BP Prop. Change (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 18/20 BP. Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(bp_brmps_diffprop)
```

##Total SCFA Concentration in Week 1 vs. Week 3 (Participants Averaged)
```{r}
#Making and Cleaning data frame with Participants by Supplement
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015")

#Filtering for Quality Compliance and Twice-Daily BRMPS Participants
participantsupps1 <- participantsupps1 %>%
  select(-starts_with("Data.consent")) %>%
  filter(Quantity.compliant == "yes") %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 18" | Supplement.consumed == "LODAAT 2xdaily 20" | Supplement.consumed == "BRMPS 2xdaily 20") %>%
  rename(Participant = Name)

#Making Week 1 Data Frame from BRMPS and Acetate
fecal1_total <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate")) %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  filter(Week == "week1") %>%
  na.omit() 

fecal1_brmps <- merge(fecal1_total,participantsupps1,by="Participant") %>%
  rename(total.wk1 = scfa_sum) %>%
  select(starts_with("Participant"),
        starts_with("total"))

#Making Week 3 Data Frame from BRMPS and Acetate
fecal3_total <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("Acetate"),
         starts_with("Butyrate"),
         starts_with("Propionate")) %>%
  mutate(scfa_sum = rowSums(.[3:5])) %>%
  filter(Week == "week3") %>%
  na.omit() 

fecal3_brmps <- merge(fecal3_total,participantsupps1,by="Participant") %>%
  rename(total.wk3 = scfa_sum) %>%
  select(starts_with("Participant"),
         starts_with("total"))

#Combining Week 1 and Week 3 BRMPS-Acetate DFs for Larger DF
total_brmps <- merge(fecal1_brmps,fecal3_brmps,by="Participant") %>%
  mutate_if(is.numeric,round)

#Appending Differences column to acetate-BRMPS df in form: week3-week1
total_brmps_diff <- total_brmps %>%
  mutate(total_diff = total.wk3 - total.wk1) %>%
  arrange(by = total_diff)

#Creating Paired Boxplot for Week 1 vs Week 3 BRMPS-Acetate Data
total_wk1_vec <- as.vector(total_brmps$total.wk1)
total_wk3_vec <- as.vector(total_brmps$total.wk3)
paired_total_df <- data.frame(Week1 = total_wk1_vec, Week3 = total_wk3_vec)
totalpairplot <- ggpaired(paired_total_df, cond1 = "Week1", cond2 = "Week3")
totalpairplot$layers <- totalpairplot$layers[-2]

#Summary Statistics and Graphics for Week 1 vs. Week 3 BRMPS-Acetate
##t-test means (wk1-wk3) difference
t.test(as.vector(total_brmps$total.wk1), 
       as.vector(total_brmps$total.wk3),
       alternative = "less",
       paired = TRUE,
       conf.level = 0.95)
summary(total_brmps_diff)
totalpairplot
totbrmps1x_barplot <- barplot(total_brmps_diff$total_diff, ylab = "Participants", xlab = "Total Conc. Change (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 18/20 and BRMPS 2x Total Change from Week 1 to Week 3", 
        horiz =  TRUE)
nrow(total_brmps_diff)
```

##pH Change in BRMPS 2x Cohort from Week 1 to Week 3
```{r}
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015") %>%
  rename(Participant = Name) %>%
  filter(Supplement.consumed == "LODAAT 2xdaily 18" | Supplement.consumed == "LODAAT 2xdaily 20" | Supplement.consumed == "BRMPS 2xdaily 20")

fullfecalsample_unaveraged<- fullfecalsample_unaveraged %>%
   mutate(scfa_sum = Acetate_mmol_kg+Butyrate_mmol_kg+Propionate_mmol_kg) 

#Making Week 1 BRMPS 2x-pH Data Frame 
ph_brmps_wk1 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week1") %>%
  na.omit() %>%
  rename(ph.wk1 = pH)

#Making Week 3 BRMPS 2x-pH Data Frame 
ph_brmps_wk3 <- fullfecalsample %>%
  select(starts_with("Participant"),
         starts_with("Week"),
         starts_with("pH")) %>%
  filter(Week == "week3") %>%
  na.omit() %>%
  rename(ph.wk3 = pH)

#Combining Week 1 and Week 3 BRMPS 2x Daily Cohort 
merged_ph_wks <- merge(ph_brmps_wk1, ph_brmps_wk3, by = "Participant") %>%
  select(-starts_with("Week"))
merged_ph_brmps<- merge(merged_ph_wks, participantsupps1, by = "Participant")
merged_ph_brmps <- merged_ph_brmps %>%
  select(starts_with("Participant"), starts_with("ph.wk")) %>%
  mutate(ph.diff = ph.wk3 - ph.wk1) %>%
  arrange(by_group = ph.diff)
#Summary Statistics and Visualizations for BRMPS 2x Daily pH Change
nrow(merged_ph_brmps) 
t.test(as.vector(merged_ph_brmps$ph.wk1), 
       as.vector(merged_ph_brmps$ph.wk3),
       alternative = "greater",
       paired = TRUE,
       conf.level = 0.95)
summary(merged_ph_brmps)
ph_brmps1x_barplot <- barplot(merged_ph_brmps$ph.diff, ylab = "Participants", 
        xlab = "pH (Week 3 - Week 1)", 
        main = "LODAAT 2x Daily 18/20 + BRMPS 2x 20 pH Change from Week 1 to Week 3", 
        horiz =  TRUE)
```
#-------------------------------------------------------------------------------
#SCFA Change vs. pH Change 
```{r}
participantsupps1 <- read.csv("./participant.csv") %>%
  filter(Semester != "Winter2015") %>%
  rename(Participant = Name) %>%
  filter(Quantity.compliant == "yes" | Quantity.compliant == "none") %>%
  select(Participant, Supplement.consumed) %>%
  na.omit()

#Isolating Week 1 Data
week1 <- fullfecalsample %>%
  filter(Week == "week1") %>%
  select(Participant, Week, pH, Acetate_mmol_kg, Butyrate_mmol_kg, 
         Propionate_mmol_kg) %>%
  mutate(scfa_sum = rowSums(.[4:6])) %>%
  na.omit()

#Isolating Week 3 Data 
week3 <- fullfecalsample %>%
  filter(Week == "week3") %>%
  select(Participant, Week, pH, Acetate_mmol_kg, Butyrate_mmol_kg, 
         Propionate_mmol_kg) %>%
  mutate(scfa_sum = rowSums(.[4:6])) %>%
  na.omit()

#Merging Week 1 and Week 3 Data and Finding pH and SCFA Differences 
allweeks <- merge(week1, week3, by = "Participant") %>%
  rename(ph.week1 = pH.x, ph.week3 = pH.y, scfa_sum.wk1 = scfa_sum.x,
         scfa_sum.wk3 = scfa_sum.y) %>%
  select(Participant, ph.week1, scfa_sum.wk1, ph.week3, scfa_sum.wk3) %>%
  mutate(ph_change = ph.week3 - ph.week1, 
         scfa_change = scfa_sum.wk3 - scfa_sum.wk1) 

allweeks <- merge(allweeks, participantsupps1, by = "Participant")

allweeks$supplement[allweeks$Supplement.consumed == "BRMPS 1xdaily 40" | allweeks$Supplement.consumed == "LODAAT 1xdaily 20"] = "RPS Once Daily"

allweeks$supplement[allweeks$Supplement.consumed == "BRMPS2xdaily 40" | allweeks$Supplement.consumed == "LODAAT 2xdaily 20" | 
allweeks$Supplement.consumed == "BRMPS 2xdaily 20" |
allweeks$Supplement.consumed == "LODAAT 2xdaily 18" |
allweeks$Supplement.consumed == "LODAAT 2xdaily 9"] = "RPS Twice Daily"

allweeks$supplement[allweeks$Supplement.consumed == "Oat 2xdaily 20" | allweeks$Supplement.consumed == "Rice 2xdaily 20"] = "Other"

allweeks$supplement[allweeks$Supplement.consumed == "none"] = "None"

allweeks <- allweeks %>%
  filter(Supplement.consumed != "transition_HiMaize 2xdaily 20",
         Supplement.consumed != "HiMaize+BRMPS 1xdaily 40")

#allweeks <- allweeks %>%
 # filter(supplement == "RPS Twice Daily")

#Creating Scatterplot for SCFA Change vs. PH Change
ggplot(allweeks, aes(x=scfa_change, y=ph_change)) +
  geom_point(alpha=0.3) +
  ggtitle("pH Change vs. SCFA Change") +
  labs(y= "pH Change (Wk3-Wk1)", x = "SCFA Change (Wk3-Wk1)") +
  theme(axis.text.x  = element_text(angle=0, hjust=0.5, size=11,colour
                                    ="black")) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(breaks= pretty_breaks()) +
  theme_bw() 

nrow(allweeks)

unique(allweeks$Supplement.consumed)

cor.test(as.vector(allweeks$scfa_change), as.vector(allweeks$ph_change), 
         alternative = "less", method = 'pearson') #cor = -0.122232 

#Scatterplot colored by Supplement
qplot(scfa_change, ph_change, data = allweeks, colour = supplement)
qplot(ph_change, scfa_change, data = allweeks, colour = supplement)
```


